%--------------------------------------------------------------------------------
%
%
%--------------------------------------------------------------------------------
\chapter*{Runtime Overview}
\addcontentsline{toc}{chapter}{Runtime Overview}

No CPU architecture with an idea of a runtime environment. Although the instruction set can be used to implement many models of runtime environments, there are common principles that exploit the CPU architecture. In fact, several CPU concepts were selected with a certain runtime already in mind. This chapter will present the Twin-64 runtime environment and describe the high level system model, the register usage convention, the execution model and calling convention.

%--------------------------------------------------------------------------------
\section*{System Environment}
\addcontentsline{toc}{section}{System Environment}

The following figure depicts a high-level overview of the overall software environment for the Twin-64 system. At the center of the execution model is the execution thread, referred to as a \textbf{task}. A task consists of the currently running execution object together with its private task-local data area, including its own stack and runtime context. All tasks of a job share a common job data area. At the highest level is the \textbf{system}, which contains the global data area for the system. 

\begin{figure}[H]
\begin{center}
\begin{tikzpicture}[scale=0.7, transform shape]
        
  	\draw[help lines, gray!70, dashed] (0,0) grid(16,6);
  	
  	\node[  tsRoundedRectangle, 
            minimum width=16cm,
            minimum height=6cm,  
            fill=white!50] at (8,3) { };
            
   	\node[  tsRoundedRectangle, 
            minimum width=6cm,
            minimum height=4cm,  
            fill=white!50] at (3.5, 2.5) { };
            
   	\node[  tsRoundedRectangle, 
            minimum width=6cm,
            minimum height=4cm,  
            fill=white!50] at (10, 2.5) { };
            
  	\filldraw (14,2.5) circle (1mm);
  	\filldraw (14.5,2.5) circle (1mm);
  	\filldraw (15,2.5) circle (1mm);
  	
  	\node[  tsRoundedRectangle, 
            minimum width=2cm,
            minimum height=2cm,  
            fill=white!50] at (2, 2) { Task };
            
  	\node[  tsRoundedRectangle, 
            minimum width=2cm,
            minimum height=2cm,  
            fill=white!50] at (4.5, 2) { Task };
  	
  	\node[  tsRoundedRectangle, 
            minimum width=2cm,
            minimum height=2cm,  
            fill=white!50] at (9, 2) { Task };
            
    \filldraw (11,2) circle (1mm);
  	\filldraw (11.5,2) circle (1mm);
  	\filldraw (12,2) circle (1mm);
  	
  	\node[	anchor=west] at ( 0.25, 5.5 ) { \LARGE System }; 
  	\node[	anchor=west] at ( 0.5, 4 ) { \LARGE Job }; 
  	\node[	anchor=west] at ( 7, 4 ) { \LARGE Job }; 
        
\end{tikzpicture}
\end{center}
\caption{System Environment}
\label{fig: System Environment}
\end{figure}

All tasks belonging to the same job share a common job data area, which is used to store information and resources that must persist for the lifetime of the job and be visible across all of its tasks. A job may represent an interactive session, such as a terminal session, or it may be a non-interactive batch operation.

The system level contains the global data area, system libraries, and all services provided by what is traditionally called the operating system. The system is responsible for creating jobs, managing their lifetime, and scheduling the tasks within them.

The number of active tasks at any moment typically corresponds to the number of available processors in the system, allowing the system to take full advantage of available parallelism. While each task maintains its own private data and execution state, tasks within a single job share access to the job data area. All tasks, regardless of job, also have access to system-level services and global resources.

%--------------------------------------------------------------------------------
\section*{Runtime Environment}
\addcontentsline{toc}{section}{Runtime Environment}

The processor is a general-purpose CPU. The runtime model presented here represents one possible implementation strategy, chosen to support the calling conventions and execution model that will be described in later sections. The following figure depicts a high-level overview of the runtime environment.

\begin{figure}[H] 
\begin{center} 
\begin{tikzpicture}[x=1cm, y=1cm, scale=0.7, transform shape] 

	\draw[help lines, gray, dashed] (0,0) grid(16,18);
	\def\dataLeft {1}; 
	\def\codeLeft {9}; 
	\def\width {4}; 
	
	%task
	\node[ tsRoundedRectangle, minimum width=16cm, minimum height=19cm, fill=white!50] at (8,8.5) { }; 
	\node[ anchor=west] at ( 0.25, 17.5 ) {  \bfseries \LARGE Task }; 
	
	%globals
	\node[ tsRectangle, minimum width=4cm, minimum height=8cm, fill=white!50, anchor=west] at ( \dataLeft, 12) { };
	\draw[thick] (\dataLeft, 13) -- (\dataLeft + \width, 13); 
	
	\node[ anchor=west] at ( \dataLeft + 0.25, 15.5 ) {  \bfseries \Large Linkage Data }; 
	\node[ anchor=west] at ( \dataLeft + 0.25, 12.75 ) {  \bfseries \Large Global Data }; 
	
	\draw[<-, thick] (\dataLeft + 4, 13) -- (\dataLeft + 5, 13); 
	\node[anchor=west] at (\dataLeft + 5.25, 13) { \Large DP }; 
	
	\node[ tsRectangle, minimum width=4cm, minimum height=1cm, fill=gray!40, anchor=west] at ( \dataLeft, 11) { }; 
	\node[anchor=west] at (\dataLeft + 0.25, 11.25) { \small Module Globals };
	
	%stack 
	\node[ tsRectangle, minimum width=4cm, minimum height=7cm, fill=white!50, anchor=west] at ( \dataLeft, 3.5) { }; 
	\node[ anchor=west] at ( \dataLeft + 0.25, 6.5 ) { \bfseries \Large Stack Data }; 
	\node[ tsRectangle, minimum width=4cm, minimum height=1cm, fill=gray!40, anchor=west] at ( \dataLeft, 3) { }; 
	\node[ anchor=west] at ( \dataLeft + 0.25, 3.125 ) { \small Stack Frame }; 
	\draw[<-, thick] (\dataLeft + 4, 2.5) -- (\dataLeft + 5, 2.5); 
	\node[anchor=west] at (\dataLeft + 5.25, 2.5) { \Large SP }; 
	
	%code 
	\node[ tsRectangle, minimum width=4cm, minimum height=10cm, fill=white!50, anchor=west] at ( \codeLeft, 11) { }; 
	\node[ anchor=west] at ( \codeLeft + 0.25, 15.5 ) {  \bfseries \Large Code }; 
	\	
	\node[ tsRectangle, minimum width=4cm, minimum height=2cm, fill=gray!40, anchor=west] at ( \codeLeft, 12) { }; 
	\node[anchor=west] at (\codeLeft + 0.25, 12.75) { \small Module Code};
	
	\node[ tsRectangle, minimum width=4cm, minimum height=0.125cm, fill=white!50, anchor=west] at ( \codeLeft, 12) { }; 
	\draw[<-, thick] ( \codeLeft + 4, 12.125 ) -- ( \codeLeft + 5, 12.125 ); 
	\node[anchor=west] at ( \codeLeft + 5.25, 12.125 ) { \Large IA }; 
	
\end{tikzpicture} 
\end{center} 
\caption{Runtime Environment} 
\label{fig: Runtime Environment} 
\end{figure}

When a compiler or assembler generates code, the fundamental unit of work is the compilation unit. A compilation unit corresponds to a single invocation of the compiler or assembler. Although syntax varies by language, the source text is organized into modules. A module is an entity that contains constant declarations, type definitions, module-level global variables, and functions.

The task’s code region contains the program text of all modules used by the task. Once loaded, this region is immutable. The currently executing instruction is identified by the instruction address (IA).

Each task also has a global data region. This task-level global data is formed by aggregating the global data sections of all modules referenced by the task. It is addressed using the data pointer (DP) plus an offset. The data region includes the task’s stack, which holds stack frames created by function calls. The active frame is located via the stack pointer (SP), and items within the stack are accessed relative to SP.

The data region further contains the linkage table, a structure that provides the information needed to access functions in external libraries. Details of this mechanism are described in the section on linkage tables.

In this runtime design, the stack grows upward: new frames occupy increasing addresses, and older frame contents are typically referenced using expressions of the form SP - <ofs>. Each stack frame includes a frame marker containing the previous SP and the return link (RL), enabling structured returns and proper unwinding of frames.

\textbf{Note.}  

%--------------------------------------------------------------------------------
\section*{Register Usage}
\addcontentsline{toc}{section}{Register Usage}

The CPU features general registers and control registers. All general registers can do computation as well as address computation. The runtime architecture builds on the general register file and assigns specific meaning to each register. 

\begin{figure}[H] 
\begin{center} 
\begin{tikzpicture}[x=1cm, y=1cm, scale=0.7, transform shape] 

    % \draw[help lines, gray, dashed] (0,0) grid(16,16); 

	% Base dimensions 
	\def\leftx{2} 
	\def\width{8} 
	\def\rightx{\leftx + \width} 
	\def\height{1} 
	
	% height of each slot % Draw outer box 
	\draw[thick] (\leftx,0) rectangle ++(\width, {16*\height}); 
	
	% Slot lines 
	\foreach \i in {1,...,15} { \draw[thick] (\leftx,{16*\height - \i*\height}) -- ++(\width,0); } 
	
	% Labels on the left side: GR0 .. GR15 
	\foreach \i in {0,...,15} { \node[anchor=east] at (\leftx - 0.2, {16*\height - \i*\height - 0.45}) {\small GR\i}; } 
	
	% Register names
	\node at (\leftx + \width/2, {16*\height - 0.45}) {Permanent zero}; 
	\node at (\leftx + \width/2, {16*\height - 1.45}) {General use / Target for ADDIL }; 
	\node at (\leftx + \width/2, {16*\height - 2.45}) {General use }; 
	\node at (\leftx + \width/2, {16*\height - 3.45}) {General use }; 
	\node at (\leftx + \width/2, {16*\height - 4.45}) {General use }; 
	\node at (\leftx + \width/2, {16*\height - 5.45}) {General use }; 
	\node at (\leftx + \width/2, {16*\height - 6.45}) {General use }; 
	\node at (\leftx + \width/2, {16*\height - 7.45}) {General use }; 
	\node at (\leftx + \width/2, {16*\height - 8.45}) {General use }; 
	\node at (\leftx + \width/2, {16*\height - 9.45}) {General use  / Arg3 }; 
	\node at (\leftx + \width/2, {16*\height - 10.45}) {General use / Arg2 }; 
	\node at (\leftx + \width/2, {16*\height - 11.45}) {General use / Arg1 / Ret1};
	\node at (\leftx + \width/2, {16*\height - 12.45}) {General use / Arg0 / Ret0}; 
	\node at (\leftx + \width/2, {16*\height - 13.45}) {General use / RL }; 
	\node at (\leftx + \width/2, {16*\height - 14.45}) {General use / SP }; 
	\node at (\leftx + \width/2, {16*\height - 15.45}) {General use / DP }; 
	
	% bracket 
	\draw[thick] [decorate, decoration={brace, amplitude=0.125cm, mirror}] (\rightx + 0.5,14) -- (\rightx + 0.5,16) node[midway, right=1cm, anchor=west]{\Large Hardware Architected }; 
	\draw[thick] [decorate, decoration={brace, amplitude=0.125cm, mirror}] (\rightx + 0.5,10) -- (\rightx + 0.5,14) node[midway, right=1cm, anchor=west]{\Large Callee Save }; 
	\draw[thick] [decorate, decoration={brace, amplitude=0.125cm, mirror}] (\rightx + 0.5,7) -- (\rightx + 0.5,10) node[midway, right=1cm, anchor=west]{\Large Caller Save }; 
	\draw[thick] [decorate, decoration={brace, amplitude=0.125cm, mirror}] (\rightx + 0.5,3) -- (\rightx + 0.5,7) node[midway, right=1cm, anchor=west]{\Large Parameter Area }; 
	\draw[thick] [decorate, decoration={brace, amplitude=0.125cm, mirror}] (\rightx + 0.5,0) -- (\rightx + 0.5,3) node[midway, right=1cm, anchor=west]{\Large Runtime Architecture }; 
	
\end{tikzpicture} 
\end{center} 
\caption{General Register Usage} 
\label{fig:general-registers-usage} 
\end{figure} 


Registers \texttt{R0} and \texttt{R1} are hardware-architected and have fixed, predefined roles within the processor. The remaining registers, \texttt{R2} through \texttt{R15}, form the general-purpose register file, whose usage is defined by the runtime environment. 

Registers \texttt{R2} through \texttt{R5} form the callee-save register set. A callee-save register is one whose value must be preserved by the called function: if the function uses such a register, it must save its original contents upon entry and restore them before returning. This ensures that long-lived values in the caller remain intact across function calls.

Registers \texttt{R6} and \texttt{R8} are designated as caller-save registers. A caller-save register may be overwritten by the called function; therefore, the caller is responsible for saving the register’s value before making a call if it must be preserved.

Argument passing is handled through registers \texttt{R9} to \texttt{R12}, known as the parameter registers. When a function is invoked, its first four arguments are placed into these registers. Registers \texttt{R11} and \texttt{R12} also serve as the return-value registers, allowing a function to return up to two values without using memory.

Control flow is supported by register \texttt{R13}, the return-link register (RL). During a function call, the branch instructions store the return address in \texttt{R13}, allowing the function to return control to its caller.

Register \texttt{R14} serves as the stack-frame pointer (SP). It marks the base of the current stack frame, which contains local variables, spilled registers, and temporaries, and supports predictable stack-based data management within the task’s runtime environment.

Finally, register \texttt{R15} is assigned as the data pointer (DP). It serves as the base reference for accessing the task’s global data region.

