%--------------------------------------------------------------------------------
%
%
%--------------------------------------------------------------------------------
\chapter*{System Organization}
\addcontentsline{toc}{chapter}{System Organization}

This chapter introduces Twin-64 by highlighting the major components and design concepts.

\section*{A Register–Memory Architecture}
\addcontentsline{toc}{section}{A Register–Memory Architecture}

Modern RISC CPUs are typically out--of--order load / store designs and feature a large number of general-purpose registers. Operations solely take place between registers. Twin-64 follows a slightly different route. It has fewer general registers, and in addition to register–-register computation, a register–-memory model is also supported. In contrast to similar historical register–-memory designs, Twin-64 has no operations that both read from and write to memory in the same instruction. For example, there is no “increment memory” instruction, since this would require two memory operations and is generally not pipeline-friendly. 

Twin-64 offers only a few addressing modes for the operand, combined with a fixed instruction word length. For most instructions one operand is a register, while the other is a flexible operand that can be an immediate, another register, or a memory address from which data is obtained or stored. The result is placed in the register, which is also the first operand. These types of machines are also called two-address machines.

\section*{System}
\addcontentsline{toc}{section}{System}

A Twin-64 system is organized around a central system bus that interconnects the key components of the architecture. Processors, main memory modules, and I/O modules are all attached as peers to this bus, enabling uniform communication and data exchange. Larger configurations may also include I/O adapters, which bridge slower peripheral buses to the central system bus while preserving a consistent programming model.

Processors initiate memory and I/O transactions via the bus and can also receive external events such as interrupts or system resets through the same interface. This modular structure provides flexibility in system design, allowing configurations to scale from simple processor–-memory systems to complex multiprocessor environments with diverse I/O subsystems.

\section*{Processor}
\addcontentsline{toc}{section}{Processor}

The Twin-64 processor implements the instruction set architecture and provides the computational resources of the system. It contains a general register set, control and status registers, and the program state register. To support virtual memory, each processor includes a Translation Lookaside Buffer (TLB) that translates virtual addresses into physical addresses. Instruction and data caches, while optional in principle, are essential for performance and tightly integrated with the memory interface. Regardless of configuration, all processor-generated addresses are virtual and must be resolved by the TLB before memory access.

\section*{Virtual Memory}
\addcontentsline{toc}{section}{Virtual Memory}

Twin-64 offers a large global address range, organized in regions. A region number and offset together form the global virtual address. regions are also associated with access rights and protection identifiers. The CPU itself can run in user or privileged mode. Twin-64 always generates virtual addresses at the CPU level. The global virtual memory range is a 52-bit space organized into a 20-bit region ID and a 32-bit byte offset. There are $2^{20}$ regions with $2^{32}$ bytes each.

Virtual addresses are translated to physical addresses by the TLB when memory is referenced. For memory management purposes, the virtual address range can also be viewed as a set of pages. A region therefore contains $2^{20}$ pages of 4 KB each. A virtual page number, combining the region and the page number within that region, is a $2^{40}$ value. Address translations are performed at the page level. 

The first regions in the virtual address space are special in that they actually represent the physical address range. An address in that range bypasses the TLB and accesses physical memory directly. 

A TLB can optionally support different page sizes in addition to the architected 4 KB page size. Supported sizes are 4 Kbyte, 64 Kbytes, 1 Mbyte and 16 MByte. It is very common for the operating system and data to be mapped in consecutive physical pages, which can then be covered by a larger page size.

\section*{Security and Protection Support}
\addcontentsline{toc}{section}{Security and Protection Support}

Twin-64 implements a protection model along two dimensions. The first dimension is \textbf{access type control} and privilege-level checking. Each page is associated with a page type. Page types include read-only, read–write, execute, and gateway. There are two privilege levels: user mode and supervisor mode. Access type and privilege level together form the access control information, which is checked for each instruction and data memory access. 

The second dimension of protection is the region ID itself. Twin-64 allows a set of region IDs to be recorded in processor control registers. For each access to a region that has the protection check bit set, one of the region ID control registers must match the region ID. If not, a protection violation trap is raised.

\section*{Privilege Changes}
\addcontentsline{toc}{section}{Privilege Changes}

Instructions can only access data or execute instructions when the privilege level matches the required privilege level. Two or four levels are the most common implementations. Changing between levels is often done with a kind of trap operation to higher-privilege software, for example the operating system kernel. However, traps are expensive, as they cause the CPU pipeline to be flushed when entering a trap handler.

Twin-64 implements gateways for privilege promotion. A special option on the unconditional branch instruction raises the privilege level when the instruction is executed on a gateway page, setting the privilege level to that of the gateway page. Returning from a higher privilege level to a code page with lower privilege level automatically resets the lower privilege level. A branch to a higher-privilege page that is not a gateway page results in a privilege violation trap.

\section*{Debugging Support}
\addcontentsline{toc}{section}{Debugging Support}

A fundamental requirement is the ability to set instruction and data breakpoints. An instruction breakpoint is triggered by encountering the \texttt{TRAP} instruction. The break instruction causes an instruction break trap and passes control to the trap handler. Since break instructions are normally not present in the instruction stream, they must be inserted dynamically in place of the instruction normally found at that address. Upon continuation, if the breakpoint is still valid, it must be reinserted after the original instruction is executed. To facilitate this mechanism, a bit in the status word provides an easy way to trap again after the instruction has been executed.

Data breakpoints are traps taken when a data reference to a page occurs. They do not modify the instruction stream. Depending on the data access, the trap may occur before or after the instruction that accesses the data. A write operation is trapped before the data is written, while a read instruction is trapped after the original instruction has executed.

\section*{Input/Output}
\addcontentsline{toc}{section}{Input/Output}

Twin-64 implements a memory-mapped I/O architecture. A portion of the physical address space is reserved for I/O modules, which respond to memory load and store instructions directed to their assigned addresses. Standard page-level protection applies during address translation, ensuring that only privileged software (typically device drivers) can safely access I/O regions. User programs may be granted direct access to I/O by mapping a user-level I/O page into virtual memory, without compromising overall system integrity.

\textbf{Direct I/O} is performed using the standard load and store instructions to access I/O modules. These operations are entirely controlled by software and may be executed in user mode if the virtual mapping allows it.

\textbf{Direct Memory Access} (DMA) modules can transfer data between memory and I/O devices independently of the processor. DMA transfers operate on physical memory, requiring that the relevant pages be locked and protected from conflicting access by the operating system. DMA transactions are initiated by issuing DMA commands, and command chaining is supported to enable complex or continuous data transfers without processor intervention.

\section*{Interrupt System}
\addcontentsline{toc}{section}{Interrupt System}

Interrupts and exceptions are events that occur asynchronously to instruction execution. Depending on the type, they occur either during the execution of an instruction or between instructions. An example of the former is an arithmetic overflow trap; an example of the latter is an external interrupt. Depending on the interrupt or exception type, the instruction is either restarted or execution continues with the next instruction.

Twin-64 implements a simple interrupt system. Each processor features an interrupt input register and an interrupt mask register. When an I/O module receives a command, it also receives information about which interrupt bit to set and which target module ID to send the interrupt to upon completion of the request. The interrupt data is compared to the interrupt mask register. If the particular bit is enabled and interrupts are globally enabled, the target processor enters the interrupt handler.

Since raising an interrupt is simply a matter of storing the interrupt request into the processor’s interrupt register, I/O modules as well as processors themselves can raise interrupts on a target processor. Processors can also send interrupts to their peers, enabling low-level processor-to-processor communication.
