\chapter*{Appendix - Instruction Commentary}
\addcontentsline{toc}{chapter}{Appendix - Instruction Commentary}

%--------------------------------------------------------------------------------
\section*{Comparison and Branch Condition Encoding}
\addcontentsline{toc}{section}{Comparison and Branch Condition Encoding}

This appendix explains the condition-code encoding and the implementation philosophy used by the \texttt{CMP}, \texttt{CBR}, \texttt{ABR} and \texttt{MBR} instructions. The design goals are to minimize hardware complexity by reusing a single subtraction/comparator path. Comparison operators are encoded in a compact and symmetric way that makes condition inversion trivial. The following instructions use the comparison operations.

\begin{itemize}
  \item \textbf{CMP} — The compare instruction uses a 2-bit \texttt{cond} field (four primitives) and produces a boolean result (0/1) in a register.
  \item \textbf{CBR} — The compare-and-branch uses a 3-bit \texttt{cond} field (eight encodings) and branches based on comparison between two registers.
  \item \textbf{ABR} — The add-and-branch performs an add operation and branches using a 3-bit \texttt{cond} field that tests the add result against zero or other simple predicates.
  \item \textbf{MBR} — move-and-branch also uses the same 3-bit \texttt{cond} field and tests the loaded result.
\end{itemize}

All three branch instructions (\texttt{CBR}, \texttt{ABR}, \texttt{MBR}) share the same 3-bit condition-code encoding so decoder logic is uniform across branch forms. \texttt{CMP} uses a compact 2-bit encoding (four primitives) for hardware efficiency; the assembler and/or an \texttt{inv} operation are used to provide the remaining relations where needed.

Branch target computation uses a sign-extended 15-bit offset shifted left by two, allowing word-aligned branch destinations within a \texttt{±64\,KB} range relative to the current instruction address. 

\subsection*{Design Rationale} 

Implementing just four compare primitives minimizes combinational logic in the arithmetic/comparator unit. The primitives chosen are \texttt{EQ}, \texttt{NE}, \texttt{LT} and \texttt{GE}. These four are symmetric in pairs: \texttt{EQ} / \texttt{NE} and \texttt{LT} / \texttt{GE}. Each pair is the logical complement of the other, so an \emph{invert} operation (flip of a single decode bit) yields the complementary condition. That property makes both ISA decoding and assembler inversion trivial.

Where a comparison such as \texttt{GT} or \texttt{LE} is required, the assembler can safely swap operands (register form, no side-effects), it will rewrite the comparison using operand swap (e.g. \texttt{A > B} → \texttt{B < A}). Where operand swapping is not possible (immediate or memory addressing modes), the condition is produced by inverting the result of an available primitive (either by an \texttt{inv} instruction that flips a 0/1 boolean, or by a branch invert bit in the branch encoding).

\subsection*{Encodings}

The compare instruction (2-bit cond) use the following compare modes:

\begin{center}
\begin{tabular}{|c|c|l|}
\hline
\textbf{Value} & \textbf{Mnemonic} & \textbf{Meaning} \\ \hline
0 & \texttt{EQ} & \texttt{op1 == op2} \\ \hline
1 & \texttt{LT} & \texttt{op1 < op2} \\ \hline
2 & \texttt{NE} & \texttt{op1 != op2} \\ \hline
3 & \texttt{GE} & \texttt{op1 >= op2} \\ \hline
\end{tabular}
\end{center}

The 2-bit encoding is chosen so that the left most bit is the \emph{invert} bit: flipping it yields the complementary condition for every pair. The other branch type instructions use the following compare modes:

\begin{center}
\begin{tabular}{|c|c|l|}
\hline
\textbf{Value} & \textbf{Mnemonic} & \textbf{Branch if} \\ \hline
0 & \texttt{EQ} & \texttt{Reg == 0} \\ \hline
1 & \texttt{LT} & \texttt{Reg < 0} \\ \hline
2 & \texttt{NE} & \texttt{Reg != 0} \\ \hline
3 & \texttt{GE} & \texttt{Reg >= 0} \\ \hline
4 & \texttt{LE} & \texttt{Reg <= 0} \\ \hline
5 & \texttt{EV} & \texttt{Reg.[0] == 0} \\ \hline
6 & \texttt{GT} & \texttt{Reg > 0} \\ \hline
7 & \texttt{OD} & \texttt{Reg.[0] != 0} \\ \hline
\end{tabular}
\end{center}

The 3-bit encoding is chosen so that middle bit is the \emph{invert} bit. Flipping it yields the complementary condition for every pair. This symmetry keeps branch inversion extremely cheap in hardware and consistent across all branch forms.

\subsection*{Logical decoding and hardware signals}

All conditions used by the comparator/branch unit are derived from a few simple signals computed by the ALU/comparator after a subtraction (or directly from the result when testing against zero). The following Verilog code sequence depicts a possible implementation.

\smallskip
\lstset{style=codesnippetstyle}
\begin{lstlisting}
wire Z  = (result == 64'd0);
wire S  = result[63];          
wire B0 = result[0];

wire cond_EQ = Z;
wire cond_LT = S;
wire cond_NE = ~Z;
wire cond_GE = ~S;
wire cond_LE = S | Z;
wire cond_GT = (~S) & (~Z);
wire cond_EV = ~B0;
wire cond_OD = B0;
\end{lstlisting}

The condition select is a multiplexer that select on of the wires above. Because all branch forms share the same condition encoding scheme, the branch unit uses a single small logic table to produce \texttt{branch\_taken}, simplifying pipeline wiring and verification.

\subsection*{CMP Instruction Assembler rules and rewrites}

If the two operands are registers it is safe to use operand swapping in order to avoid generating extra instructions (invert or boolean complement) and to keep assembly readable. When a programmer writes a relation that is not one of the four primitives, the assembler rewrites it to:
\smallskip
\lstset{style=codesnippetstyle}
\begin{lstlisting}
CMP.GT rR, rA, rB -> CMP.LT rR, rB, rA
CMP.LE rR, rA, rB -> CMP.GE rR, rB, rA
\end{lstlisting}
Swapping is only applied when both operands are registers. When the second operand is an immediate or a memory reference, swapping is not possible. In that case the assembler emits the corresponding primitive followed by an \texttt{invert} operation:
\smallskip
\lstset{style=codesnippetstyle}
\begin{lstlisting}
CMP.GT rR, rA, <arg> -> CMP.GE rR, rA, <arg> ; OR.N rR, rR, R0
CMP.LE rR, rA, <arg> -> CMP.LT rR, rA, <arg> ; OR.N rR, rR, R0
\end{lstlisting}
When generating conditional code, the assembler may omit the explicit invert operation if the logic of the surrounding construct (for example, an \texttt{IF} statement) allows the THEN and ELSE branches to be swapped without changing program semantics.
