%--------------------------------------------------------------------------------
\section*{Boolean Operation Instructions}
\addcontentsline{toc}{section}{Boolean Operation Instructions}

The Twin-64 architecture supports the standard Boolean instructions: 
\texttt{AND}, \texttt{OR}, and \texttt{XOR}. These instructions allow 
complementing the second operand as well as the output, enabling flexible 
masking and bit-field manipulation.

For example, one can apply a mask directly or use its complement to isolate 
or clear specific bits in a register efficiently.

When a Boolean instruction uses an immediate value as one operand, there are 
two possible interpretations for the immediate field:
\begin{itemize}
    \item \textbf{Unsigned zero-extended:} the immediate is treated as a 
          positive value and zero-extended to the machine word size. This 
          is convenient for constructing masks affecting only low-order bits.
    \item \textbf{Signed sign-extended:} the immediate is interpreted as a 
          signed value and extended with its sign bit. This allows the use 
          of negative constants, which produce masks with high-order bits set.
\end{itemize}

Twin-64 adopts \emph{sign-extension} for logical immediate values. The primary
motivation is uniformity: the logical I-type instructions share the same 
immediate decoding path as arithmetic instructions such as \texttt{ADDI}, 
avoiding additional decode logic.  

Sign-extension also enables several useful bit-manipulation idioms without 
requiring extra instructions. Examples include:
\begin{itemize}
    \item \texttt{AND Rn, Rm, -2} produces the mask \texttt{0x\ldots FFFE}, 
          which clears the least significant bit.
    \item \texttt{AND Rn, Rm, -4096} yields \texttt{0x\ldots F800}, aligning 
          an address to a 4\,KiB boundary.
    \item \texttt{XOR Rn, Rm, -1} results in \texttt{0x\ldots FFFF}, effectively 
          computing a bitwise NOT in a single instruction.
    \item \texttt{OR Rn, Rm, -256} creates \texttt{0x\ldots FF00}, forcing 
          the upper byte of a register to ones.
\end{itemize}

Using sign-extension for logical immediate values therefore provides both a 
straightforward hardware implementation and a flexible set of common masking 
and bit-twiddling operations.

%--------------------------------------------------------------------------------
\section*{Bit Operation Instructions}
\addcontentsline{toc}{section}{Bit Operation Instructions}

Twin-64 provides specialized bit-manipulation instructions such as 
\texttt{EXTR}, \texttt{DEP}, and \texttt{DSR} to support shifts, rotates, 
and other fine-grained operations on register fields. 

These instructions are commonly used for:
\begin{itemize}
    \item Extracting a subset of bits from a register (\texttt{EXTR}).
    \item Depositing a value into a specific bit field (\texttt{DEP}).
    \item Performing data shifts and rotations (\texttt{DSR}), which can emulate 
          traditional shift/rotate instructions without dedicated opcodes.
\end{itemize}

For example, the \texttt{EXTR} instruction can be combined with 
\texttt{AND} masks to isolate or manipulate specific bits efficiently. This 
approach provides the functionality of classic shift and rotate instructions, 
while maintaining a compact instruction set.




