%--------------------------------------------------------------------------------
%
%
%--------------------------------------------------------------------------------
\pagebreak
\section*{WF — Window forward}
\addcontentsline{toc}{section}{WF — Window forward}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Syntax}
\texttt{wf [ <amt> ] [ , <winNum> ]} \\
\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Description}

The \texttt{WF} command scrolls a window forward.
The optional \texttt{<amt>} parameter specifies the number of items to advance.
The meaning of an item is window-type specific.
For example, a physical memory window advances in units of 8 bytes, while a cache
window advances in units of cache lines.

If the \texttt{<amt>} parameter is omitted, the number of items to advance is
derived from the number of items per line and the number of lines currently
displayed in the window. For example, a cache window displaying three lines
advances by three cache-line entries.

Each scrollable window has an item address limit determined by its window type.
When scrolling past this limit, the address is set to the limit minus the number of lines to show.

If the window number parameter is omitted, the command applies to the current
window. Otherwise, it applies to the specified window.

\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescOpEntry}{Example}
\begin{lstlisting}[style=iPageOpStyle]
wn tlb, 3           # Create a TLB window for module 3.
wf                  # Advance the window by one page.
wf 5                # Advance five items in the current window.

wn mem, 0x1000      # Create a memory window.
wf                  # Window advances to address 0x1080.
\end{lstlisting}
\end{iPageDescOpEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Notes}
None.
\end{iPageDescEntry}
