%--------------------------------------------------------------------------------
%
%
%--------------------------------------------------------------------------------
\pagebreak
\section*{WB — Window backward}
\addcontentsline{toc}{section}{WB — Window backward}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Syntax}
\texttt{wb [ <amt> ] [ , <winNum> ]} \\
\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Description}

The \texttt{WB} command scrolls a window backward. The optional \texttt{<amt>} parameter specifies the number of items to move backward. The meaning of an item is window-type specific. For example, a physical memory window moves in units of 8 bytes, while a cache window moves in units of cache lines. If this parameter is omitted, the number of items to move backward is derived from the number of items per line and the number of lines currently displayed in the window. For example, a cache window displaying three lines moves backward by three cache-line entries.

Each scrollable window has an item address limit determined by its window type. When scrolling past this limit, the address is set to the limit minus the number of lines to show.

If the window number parameter is omitted, the command applies to the current window. Otherwise, it applies to the specified window.
\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescOpEntry}{Example}
\begin{lstlisting}[style=iPageOpStyle]
wn icache, 3        # Create a cache window for module 3.
wj 10               # set the window content start to entry 10.
wb 5                # Move backward five items.

wn mem, 0x1000      # Create a memory window.
wb                  # Window moves backward to address 0x0F80.
\end{lstlisting}
\end{iPageDescOpEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Notes}
None.
\end{iPageDescEntry}
