%--------------------------------------------------------------------------------
%
%
%--------------------------------------------------------------------------------
\pagebreak
\section*{WN — Create a window}
\addcontentsline{toc}{section}{WN — Create a window}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Syntax}
\texttt{wn <type> [ , <arg1> [ , <arg2> ]] } \\
\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Description}

The \texttt{WN} command creates a new window. The mandatory \texttt{<type>}
parameter specifies the type of window to create. Optional arguments provide
additional data required for the window creation.

The newly created window becomes the current window and is initialized with
default attributes, such as the number of lines, layout, and numeric radix.

\begin{flushleft}
\begin{tabularx}{\textwidth}{|l|X|X|p{5cm}|}
\hline
\textbf{Type} & \textbf{Arg 1} & \textbf{Arg 2} & \textbf{Action} \\ \hline
\texttt{MEM}         & \texttt{memAdr} & \texttt{-} & Create a memory window. \\ \hline
\texttt{PROC}        & \texttt{modNum} & \texttt{-} & Create all PROC-related windows (CPU, TLB) for the module. \\ \hline
\texttt{CPU}         & \texttt{modNum} & \texttt{-} & Create a CPU window. \\ \hline
\texttt{TLB}         & \texttt{modNum} & \texttt{-} & Create a TLB window. \\ \hline
\texttt{TEXT}        & \texttt{filePath} & \texttt{-} & Create a text window displaying the content of a text file. \\ \hline
\texttt{CODE}        & \texttt{memAdr} & \texttt{-} & Create a memory window showing assembly code. The first argument is the memory address, the second the module number of the processor.\\ \hline
\end{tabularx}
\end{flushleft}

The window type \texttt{PROC} is a collection that includes CPU and TLB  windows for the specified module.

The window type \texttt{CODE} accepts in addition to the memory address a processor module number. When the module is known, the module instruction address can be compared to the memory window address range and an indicator where the instruction is will be displayed.

\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescOpEntry}{Example}
\begin{lstlisting}[style=iPageOpStyle]
wn mem, 0x1000      # Create a memory window at address 0x1000.
wn cpu, 3           # Create a CPU window for module 3.
wn tlb, 4           # Create a TLB window for module 4.
\end{lstlisting}
\end{iPageDescOpEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Notes}
All newly created windows start with default attributes, which can be modified
later using commands like \texttt{WL} (set window lines) or \texttt{WR} (set window radix), if applicable for the window.
\end{iPageDescEntry}
