%--------------------------------------------------------------------------------
%
%
%--------------------------------------------------------------------------------
\pagebreak
\section*{WS — Move windows to stack}
\addcontentsline{toc}{section}{WS — Move windows to stack}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Syntax}
\texttt{ws <stackNum> [ , <wStart> ] [ , <wEnd> ]} \\
\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Description}

The \texttt{WS} command moves one or more windows into a specific stack.  
A stack is a vertical column of windows. Multiple stacks are arranged
horizontally on the screen and are ordered by their stack number.

Within a stack, windows are sorted by their window number.  
Windows can be enabled, disabled, or hidden independently of their stack
placement.

If the stackNum parameter is omitted, the current stack is taken as the stack number argument. 

If the window range is omitted, the current window is selected. If just the start window is specified, this window is selected for moving to the stack.

\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescOpEntry}{Example}
\begin{lstlisting}[style=iPageOpStyle]
wn mem, 0x1000       # Create a memory window, window 1.
wn proc, 3            # Create processor windows, window 2 to 6.
ws 2, 2, 6            # Move windows 2 through 6 to stack 2.
ws 1, 1               # Move window 1 to stack 1.
\end{lstlisting}
\end{iPageDescOpEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Notes}
Stacks allow grouping windows for easier organization on the screen.  
Windows maintain all their attributes when moved between stacks.
\end{iPageDescEntry}
