%--------------------------------------------------------------------------------
%
%
%--------------------------------------------------------------------------------
\pagebreak

\section*{LOG - Write a log entry to the log file}
\addcontentsline{toc}{section}{LOG - Write a log entry to the log file}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Syntax}
\texttt{log  <logMsg>  [ , <expr> ]}
\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Description}

The \texttt{LOG} command writes an entry to the log file. A log entry consists of a message string \texttt{logMsg} and an optional expression value. The expression is evaluated and its resulting value is included in the log output. If no log file is available, the command report an error.

The \texttt{LOG} command is intended for recording diagnostic information, execution traces, and test-related output during script execution. It can be used independently, or in combination with the \texttt{CHECK} and \texttt{ASSERT} commands.

Both \texttt{CHECK} and \texttt{ASSERT} automatically write their results to the log file when logging is enabled. This includes the evaluation status (\texttt{PASS}/\texttt{FAIL}) and any associated message string.

\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Example}

The following examples demonstrate the usage of the \texttt{LOG} command. The first example logs a simple message, while the second logs a message together with the result of a boolean expression.

\begin{lstlisting}[style=iPageOpStyle]

(4) -> log "a log entry message"

(5) -> log "test a boolean expression", 2 < 3

(6) ->

\end{lstlisting}
\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Notes}

Log files can be used to record execution traces and test results for debugging and validation purposes. The \texttt{CHECK} and \texttt{ASSERT} commands automatically extend this log by recording their evaluation results when logging is enabled.

\texttt{LOG} does not affect program flow and has no side effects other than producing log output.

\end{iPageDescEntry}