\chapter*{Simulator Line Commands}
\addcontentsline{toc}{chapter}{Simulator Line Commands}

This chapter presents the line mode commands. Although they are called line mode commands, these commands are also available when windows are enabled. The following table is summary of the available commands. Like all commands, they are entered in the command window.

\section*{Command Summary}
\addcontentsline{toc}{section}{Command Summary}

\begin{longtable}{@{}p{0.3\linewidth}p{0.6\linewidth}@{}}
    \caption{Simulator Line Commands} \\
    \toprule
    \textbf{Command} & \textbf{Purpose} \\
    \midrule
    \endfirsthead
    \toprule
     \textbf{Command} & \textbf{Purpose} \\
    \midrule
    \endhead
    \midrule
    \multicolumn{2}{r}{\textit{Continued on next page}} \\
    \midrule
    \endfoot
    \bottomrule
    \endlastfoot
    \textbf{ASSERT} & ensure condition \\
	\midrule
	\textbf{CHECK} & test condition \\
	\midrule
    \textbf{DM} & display memory \\
	\midrule
	\textbf{DMOD} & display configured Twin-64 modules \\
	\midrule
	\textbf{DWIN} & display window list \\
	\midrule
	\textbf{DO} & execute a command from the command history stack \\
	\midrule
	\textbf{DW} & display windows \\
	\midrule
	\textbf{ENV} & display, add, modify environment variables \\
	\midrule
	\textbf{EXIT} & leave the Twin-64 simulator \\
	\midrule
	\textbf{HALT} & halt a module \\
	\midrule
	\textbf{HELP} & display help information \\
	\midrule
	\textbf{HIST} & display command history stack \\
	\midrule
	\textbf{ITLB} & insert into the unified TLB of a CPU \\
	\midrule
	\textbf{LOADELF} & load ELF file into memory \\
	\midrule
	\textbf{MB} & modify memory byte data item ( 8 bit )\\
	\midrule
	\textbf{MS} & modify memory short data item ( 16 bit )\\
	\midrule
	\textbf{MW} & modify memory word data item ( 32 bit )\\
	\midrule
	\textbf{MD} & modify memory double data item ( 64 bit )\\
	\midrule
	\textbf{MR} & modify a processor register \\
	\midrule
	\textbf{NMOD} & add a module to the simulated environment \\
	\midrule
	\textbf{PTLB} & purge from the unified TLB of a CPU \\
	\midrule
	\textbf{REDO} & edit and do a command from the command history stack \\
	\midrule
	\textbf{RMOD} & remove a module from the simulated environment \\
	\midrule
	\textbf{RESET} & reset a module \\
	\midrule
	\textbf{RUN} & run a configured simulation \\
	\midrule
	\textbf{STEP} & step through a simulation\\
	\midrule
	\textbf{W} & evaluate and display an expression \\
	\midrule
	\textbf{XF} & execute commands from a file \\
	\end{longtable}
	

\section*{Command Syntax}
\addcontentsline{toc}{section}{Command Syntax}

Command syntax is described using the following conventions. Reserved words
are written in uppercase letters. Parameters are written in lowercase letters.

Optional elements are enclosed in square brackets \texttt{[ ]}. Elements that
may be repeated are enclosed in curly braces \texttt{\{ \}}.

Numeric values are written in decimal notation by default. When expressed in
hexadecimal notation, numeric values are prefixed with \texttt{0x}. For improved
readability, underscore characters may be used as digit separators within
hexadecimal values.

String literals are enclosed in double quotation marks (\texttt{" "}). Longer
strings may be constructed by placing multiple string literals adjacent to
each other.

A command line may extend across multiple lines. To continue a command on the
next line, place a backslash (\texttt{\textbackslash}) at the end of the line
before the carriage return.

The following sections describe each command in detail, including its purpose,
syntax, parameters, and usage examples.

\input{chapter-simulator-line-commands/chapter-sim-cmd-ASSERT}
\input{chapter-simulator-line-commands/chapter-sim-cmd-CHECK}
\input{chapter-simulator-line-commands/chapter-sim-cmd-DM}
\input{chapter-simulator-line-commands/chapter-sim-cmd-DMOD}
\input{chapter-simulator-line-commands/chapter-sim-cmd-DO}
\input{chapter-simulator-line-commands/chapter-sim-cmd-DWIN}
\input{chapter-simulator-line-commands/chapter-sim-cmd-ENV}
\input{chapter-simulator-line-commands/chapter-sim-cmd-EXIT}
\input{chapter-simulator-line-commands/chapter-sim-cmd-HALT}
\input{chapter-simulator-line-commands/chapter-sim-cmd-HELP}
\input{chapter-simulator-line-commands/chapter-sim-cmd-HIST}
\input{chapter-simulator-line-commands/chapter-sim-cmd-ITLB}
\input{chapter-simulator-line-commands/chapter-sim-cmd-LOADELF}
\input{chapter-simulator-line-commands/chapter-sim-cmd-MM}
\input{chapter-simulator-line-commands/chapter-sim-cmd-MR}
\input{chapter-simulator-line-commands/chapter-sim-cmd-NMOD}
\input{chapter-simulator-line-commands/chapter-sim-cmd-PTLB}
\input{chapter-simulator-line-commands/chapter-sim-cmd-REDO}
\input{chapter-simulator-line-commands/chapter-sim-cmd-RMOD}
\input{chapter-simulator-line-commands/chapter-sim-cmd-RESET}
\input{chapter-simulator-line-commands/chapter-sim-cmd-RUN}
\input{chapter-simulator-line-commands/chapter-sim-cmd-STEP}
\input{chapter-simulator-line-commands/chapter-sim-cmd-W}
\input{chapter-simulator-line-commands/chapter-sim-cmd-XF}
