%--------------------------------------------------------------------------------
%
%
%--------------------------------------------------------------------------------
\pagebreak
\section*{IITLB — Insert into Instruction TLB}
\addcontentsline{toc}{section}{IITLB — Insert into Instruction TLB}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Syntax}
\texttt{iitlb <vAdr> , <pAdr> , <size> , <acc> [ , L [ , U ]]}
\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Description}

The \texttt{IITLB} command inserts a translation entry into the instruction
Translation Lookaside Buffer (TLB). The \texttt{vAdr} parameter specifies the
virtual address of the mapping, while \texttt{pAdr} specifies the corresponding
physical memory address.

To reduce TLB lookup overhead and improve translation efficiency, the instruction
TLB supports mappings over contiguous virtual address ranges rather than only
fixed-size pages. Multiple page sizes are supported, allowing software to choose
an appropriate granularity based on usage patterns. The supported page sizes are
4~KB, 64~KB, 1~MB, and 16~MB.

The \texttt{acc} parameter encodes the access permissions and privilege
requirements for the translation. It specifies whether the mapped page may be read, written, or executed, as well as whether it is a gateway page along with the minimum privilege level required for instruction fetch. Every instruction fetch is checked against these access attributes, and violations result in a protection fault.

Optional flags may be supplied to control the behavior of the TLB entry. These
flags determine how the entry participates in replacement and caching decisions.

\begin{smallGapItemize}
  \item \textbf{L} — Locks the TLB entry so that it is not considered during
        replacement candidate selection.
  \item \textbf{U} — Marks the mapped page as uncached; instruction fetches bypass
        the instruction cache for this translation.
\end{smallGapItemize}

\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescOpEntry}{Example}
\begin{lstlisting}[style=iPageOpStyle]

# insert a translation for 0x1000_FFFF_0000 at 0x5000 with a 
# page size of 4Kb, access rights read-write, supervisor level
# and lock the translation 

(3)->iitlb 0x1000_FFFF_0000, 0x5000, 0, 0x4, L

\end{lstlisting}
\end{iPageDescOpEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Notes}
None.
\end{iPageDescEntry}
