%--------------------------------------------------------------------------------
%
%
%--------------------------------------------------------------------------------
\pagebreak
\section*{CHECK - Test a condition}
\addcontentsline{toc}{section}{CHECK - Test a condition}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Syntax}
\texttt{check  <expr>  [ , <msgStr> ]}
\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Description}

The \texttt{CHECK} command evaluates the specified expression and reports whether the condition passes or fails. Unlike the \texttt{ASSERT} command, a failed check does not generate an error or interrupt execution. Instead, the result is reported and execution continues normally.

If the expression evaluates to true, the command reports a \texttt{PASS} condition. If the expression evaluates to false, it reports a \texttt{FAIL} condition. An optional message string may be provided to describe the purpose of the check or to identify the condition being tested. If the message text is omitted, the default text is taken from the environment variable \texttt{ENV\_CHECK\_DEF\_MSG}.

If a log file is available, the \texttt{CHECK} command also writes its result to the log file, including the evaluation status and any associated message.

The \texttt{CHECK} command is intended for lightweight testing, validation, and interactive diagnostics where failures should be reported without stopping execution.

\end{iPageDescEntry}


%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Example}

The following examples demonstrate the \texttt{CHECK} command. The first condition passes, while the second condition fails and displays the associated message string.

\begin{lstlisting}[style=iPageOpStyle]
(4) -> check ( 2 < 3 )
PASS
(5) -> check ( 3 < 2 ), "A simple test"
"A simple test" : FAIL
(6) ->
\end{lstlisting}
\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Notes}

\texttt{CHECK} reports failures but does not stop execution. The optional message string can be used to identify individual tests. The expression is considered successful if it evaluates to a non-zero or true value.

\end{iPageDescEntry}