%--------------------------------------------------------------------------------
%
%
%--------------------------------------------------------------------------------
\chapter*{Processing Resources}
\addcontentsline{toc}{chapter}{Processing Resources}

This chapter describes the processing resources and data types in a Twin-64 system. Twin-64 provides a set of general-purpose registers, a set of control registers, and the Program State Word (PSW), which contains the current instruction address and processor status. 
\begin{figure}[H]
\begin{center}
\begin{tikzpicture}[scale=0.8, transform shape]

    % \draw[help lines, gray!70, dashed] (0,0) grid(16,4);
    \def\leftofs{1}

    % --- General Registers -----------------------------------------
    \node[tsRectangle,
          minimum width=4cm,
          minimum height=4cm,
          fill=white,
          anchor=west] (genreg) at (\leftofs,2)  {General\\Registers};

    % 16 subdivisions (4cm tall → 0.25cm per row)
    \foreach \i in {1,...,15} {
        \draw[gray!60]
            ([yshift=2cm-0.25cm*\i]genreg.west) --
            ([yshift=2cm-0.25cm*\i]genreg.east);
    }


    % --- Control Registers -----------------------------------------
    \node[tsRectangle,
          minimum width=4cm,
          minimum height=4cm,
          fill=white,
          anchor=west] (ctrlreg) at (\leftofs+5,2) {Control\\Registers};

    \foreach \i in {1,...,15} {
        \draw[gray!60]
            ([yshift=2cm-0.25cm*\i]ctrlreg.west) --
            ([yshift=2cm-0.25cm*\i]ctrlreg.east);
    }

    % --- Instruction Address block ---------------------------------
    \node[tsRectangle,
          minimum width=4cm,
          minimum height=0.5cm,
          fill=white,
          anchor=west] at (\leftofs + 10,3.75) { Instruction Address };

    % --- Status block ----------------------------------------------
    \node[tsRectangle,
          minimum width=1.5cm,
          minimum height=0.5cm,
          fill=white,
          anchor=west] at (\leftofs + 10, 3) { Status };

\end{tikzpicture}
\end{center}
\caption{Software Accessible Registers}
\label{fig:Software Accessible Registers}
\end{figure}

General registers are used for computation and address storage, control registers manage processor state, and the PSW holds the current instruction address along with the processor status.

\section*{General Registers}
\addcontentsline{toc}{section}{General Registers}

Twin-64 features a set of 16 general purposes registers. They are used for all computations including address arithmetic. 

\begin{figure}[H]
\begin{center}
\begin{tikzpicture}[x=1cm, y=1cm, scale=0.7, transform shape]
        
  	% \draw[help lines, gray!70, dashed] (0,0) grid(12,9);

	% Rectangle dimensions
	\def\width{10}
	\def\leftofs{2}
	\def\rightofs{\leftofs + \width}
	\def\middleofs{ \leftofs + \width / 2 }
	
	% Bits
	\node at (\leftofs + 0.25,8 + 0.25) {\small 63};
	\node at (\rightofs - 0.25,8 + 0.25) {\small 0};

	% Box
	\draw[thick] (\leftofs, 0) -- (\leftofs, 8);
	\draw[thick] (\rightofs, 0) -- (\rightofs, 8);
	\draw[thick] (\leftofs, 0) -- (\rightofs, 0);
	\draw[thick] (\leftofs, 8) -- (\rightofs, 8);
	  		
	% Fields  		
  	\draw[thick] (\leftofs, 7) -- (\rightofs, 7);
  	\node at ({\leftofs - 0.75}, {7 + 0.5}) { GR0 };
  	\node at (\middleofs, {7 + 0.5}) { Permanent zero };
  		
  	\draw[thick] (\leftofs, 6) -- (\rightofs, 6);
  	\node at ({\leftofs - 0.75}, {6 + 0.5}) { GR1};
  	\node at (\middleofs, {6 + 0.5}) { Target for ADDIL or General use };	
  		
  	\draw[thick] (\leftofs, 5) -- (\rightofs, 5);
  	\node at ({\leftofs - 0.75}, {5 + 0.5}) { GR2};
  	\node at (\middleofs, {5 + 0.5}) { General use };	
  		
  	\node at (\middleofs, {3 + 0.5}) { \huge{...}};
  	
  	\draw[thick] (\leftofs, 2) -- (\rightofs, 2);
  	\node at ({\leftofs - 0.75}, {1 + 0.5}) { GR14};
  	\node at (\middleofs, {1 + 0.5}) { General use };	
  		
  	\draw[thick] (\leftofs, 1) -- (\rightofs, 1);
  	\node at ({\leftofs - 0.75}, {0 + 0.5}) { GR15};
  	\node at (\middleofs, {0 + 0.5}) { General use };	
  
\end{tikzpicture}
\end{center}
\caption{General registers}
\label{fig:General registers}
\end{figure}

Some general registers have dedicated purposes. Register zero always returns a zero value when read, and writing to it has no effect. Although the CPU has only 16 general registers, implementing such a register greatly simplifies instruction design, for example by providing a target when the result is not needed. General register one is a scratch register and also serves as an implicit target for certain instructions. Other general registers may have dedicated roles as defined by the runtime architecture conventions.

%--------------------------------------------------------------------------------
\section*{Control Registers}
\addcontentsline{toc}{section}{Control Registers}

Twin-64 has 16 defined control registers, numbered \texttt{CR0} to \texttt{CR15}. Two instructions allow moving data between a control register and a general register in either direction. A move to a control register always transfers the full 64-bit word. A move from a control register to a general register transfers the value right-aligned to the size of the control register. The following figure depicts the control registers defined for Twin-64.

\begin{figure}[H]
\begin{center}
\begin{tikzpicture}[x=1cm, y=1cm, scale=0.7, transform shape]
        
  	% \draw[help lines, gray!70, dashed] (0,0) grid(12,17);
  	
  	% Rectangle dimensions
	\def\width{10}
	\def\leftofs{2}
	\def\rightofs{\leftofs + \width}
  	\def\middleofs{ \leftofs + \width / 2 }
  	
  	% Bits
  	\node at (\leftofs + 0.25,16 + 0.25) {\small 63};
  	\node at (\middleofs +0.25,16 + 0.25) {\small 32};
	\node at (\middleofs - 0.25,16 + 0.25) {\small 31};
	\node at (\rightofs - 0.25,16 + 0.25) {\small 0};

	% Box
	\draw[thick] (\leftofs, 0) -- (\leftofs, 16);
	\draw[thick] (\rightofs, 0) -- (\rightofs, 16);
	\draw[thick] (\leftofs, 0) -- (\rightofs, 0);
	\draw[thick] (\leftofs, 16) -- (\rightofs, 16);
	
	% Fields 
	% CR0
	\draw[thick] (\leftofs, 15) -- (\rightofs, 15);
  	\node at ({\leftofs - 0.75}, {15 + 0.5}) { CR0 };
  	\node at ( \middleofs, {15 + 0.5}) { Processor configuration (PCR)};	
  	
  	% CR1
  	\draw[thick] (\leftofs, 14) -- (\rightofs, 14);
  	\node at ({\leftofs - 0.75}, {14 + 0.5}) { CR1 };
  	\node at ( \middleofs, {14 + 0.5}) { Recovery Counter (RCTR)};
  	
  	% CR2
  	\draw[thick] (\leftofs, 13) -- (\rightofs, 13);
  	\draw[thick] (\rightofs - 1.5, 14) -- (\rightofs - 1.5, 13);
  	\node at ({\leftofs - 0.75}, {13 + 0.5}) { CR2 };
  	\node at ( \leftofs + 3.5, {13 + 0.5}) { Reserved };
  	\node at ( \rightofs - 0.75, {13 + 0.5}) { SAR };
  	
  	% CR3
  	\draw[thick] (\leftofs, 12) -- (\rightofs, 12);
  	\node at ({\leftofs - 0.75}, {12 + 0.5}) { CR3 };
  	\node at ( \middleofs, {12 + 0.5}) { Reserved };
  	
  	% CR4
  	\draw[thick] (\leftofs, 11) -- (\rightofs, 11);
  	\draw[thick] (\leftofs + \width / 2, 11) -- (\leftofs + \width / 2, 12);
  	\draw[thick] (\leftofs + \width / 2 - 0.5, 11) -- (\leftofs + \width / 2 - 0.5, 12);
  	\draw[thick] (\rightofs - 0.5, 11) -- (\rightofs - 0.5, 12);
  	\node at ({\leftofs - 0.75}, {11 + 0.5}) { CR4 };
  	\node at ({\leftofs + 2.25}, {11 + 0.5}) { Region Id 1 (RID1)};
  	\node at ({\leftofs + \width/2 - 0.25}, {11 + 0.5}) { W };
  	\node at ({\rightofs - 2.75}, {11 + 0.5}) { Region Id 2 (RID2)};
  	\node at ({\rightofs - 0.25}, {11 + 0.5}) { W };
  	
  	% CR5
  	\draw[thick] (\leftofs, 10) -- (\rightofs, 10);
  	\draw[thick] (\leftofs + \width / 2, 10) -- (\leftofs + \width / 2, 11);
  	\draw[thick] (\leftofs + \width / 2 - 0.5, 10) -- (\leftofs + \width / 2 - 0.5, 11);
  	\draw[thick] (\rightofs - 0.5, 10) -- (\rightofs - 0.5, 11);
  	\node at ({\leftofs - 0.75}, {10 + 0.5}) { CR5 };
  	\node at ({\leftofs + 2.25}, {10 + 0.5}) { Region Id 3 (RID3)};
  	\node at ({\leftofs + \width/2 - 0.25}, {10 + 0.5}) { W };
  	\node at ({\rightofs - 2.75}, {10 + 0.5}) { Region Id 4 (RID4)};
  	\node at ({\rightofs - 0.25}, {10 + 0.5}) { W };
  	
  	%CR6
  	\draw[thick] (\leftofs, 9) -- (\rightofs, 9);
  	\draw[thick] (\leftofs + \width / 2, 9) -- (\leftofs + \width / 2, 10);
  	\draw[thick] (\leftofs + \width / 2 - 0.5, 9) -- (\leftofs + \width / 2 - 0.5, 10);
  	\draw[thick] (\rightofs - 0.5, 9) -- (\rightofs - 0.5, 10);
  	\node at ({\leftofs - 0.75}, {9 + 0.5}) { CR6 };
  	\node at ({\leftofs + 2.25}, {9 + 0.5}) { Region Id 5 (RID5)};
  	\node at ({\leftofs + \width/2 - 0.25}, {9 + 0.5}) { W };
  	\node at ({\rightofs - 2.75}, {9 + 0.5}) { Region Id 6 (RID6)};
  	\node at ({\rightofs - 0.25}, {9 + 0.5}) { W };
  	
  	% CR7
  	\draw[thick] (\leftofs, 8) -- (\rightofs, 8);
  	\draw[thick] (\leftofs + \width / 2, 8) -- (\leftofs + \width / 2, 9);
  	\draw[thick] (\leftofs + \width / 2 - 0.5, 8) -- (\leftofs + \width / 2 - 0.5, 9);
  	\draw[thick] (\rightofs - 0.5, 8) -- (\rightofs - 0.5, 9);
  	\node at ({\leftofs - 0.75}, {8 + 0.5}) { CR7 };
  	\node at ({\leftofs + 2.25}, {8 + 0.5}) { Region Id 7 (RID7)};
  	\node at ({\leftofs + \width/2 - 0.25}, {8 + 0.5}) { W };
  	\node at ({\rightofs - 2.75}, {8 + 0.5}) { Region Id 8 (RID8)};
  	\node at ({\rightofs - 0.25}, {8 + 0.5}) { W };
  	
  	% CR8
  	\draw[thick] (\leftofs, 7) -- (\rightofs, 7);
  	\node at ({\leftofs - 0.75}, {7 + 0.5}) { CR8 };
  	\node at ( \middleofs, {7 + 0.5}) { Interrupt Vector Table Address (IVA) };
  	
  	% CR9
  	\draw[thick] (\leftofs, 6) -- (\rightofs, 6);
  	\node at ({\leftofs - 0.75}, {6 + 0.5}) { CR9 };
  	\node at ( \middleofs, {6 + 0.5}) { External Interrupt Enable Mask (EIEM)};
  	
  	% CR10
  	\draw[thick] (\leftofs, 5) -- (\rightofs, 5);
  	\node at ({\leftofs - 0.75}, {5 + 0.5}) { CR10 };
  	\node at ( \middleofs, {5 + 0.5}) { Interruption PSW (IPSW) };
  	
  	% CR11
  	\draw[thick] (\leftofs, 4) -- (\rightofs, 4);
  	\node at ({\leftofs - 0.75}, {4 + 0.5}) { CR11 };
  	\node at ( \middleofs, {4 + 0.5}) { Interruption Instruction (IARG0)};
  	
  	% CR12
  	\draw[thick] (\leftofs, 3) -- (\rightofs, 3);
  	\node at ({\leftofs - 0.75}, {3 + 0.5}) { CR12 };
  	\node at ( \middleofs, {3 + 0.5}) { Interruption Address (IARG1)};
  	
  	% CR13
  	\draw[thick] (\leftofs, 2) -- (\rightofs, 2);
  	\node at ({\leftofs - 0.75}, {2 + 0.5}) { CR13 };
  	\node at ( \middleofs, {2 + 0.5}) { Scratch Reg 0 };

  	% CR14
  	\draw[thick] (\leftofs, 1) -- (\rightofs, 1);;
  	\node at ({\leftofs - 0.75}, {1 + 0.5}) { CR14 };
  	\node at ( \middleofs, {1 + 0.5}) { Scratch Reg 1 };
  	
  	% CR15
  	\draw[thick] (\leftofs, 0) -- (\rightofs, 0);
  	\node at ({\leftofs - 0.75}, {0 + 0.5}) { CR15 };
  	\node at ( \middleofs, {0 + 0.5}) { Scratch Reg 2 };
			
\end{tikzpicture}
\end{center}
\caption{Control registers}
\label{fig:Control registers}
\end{figure}

\textbf{Note:} We may need more than 16 control registers. Under evaluation.

\subsection*{Processor Configuration Register}
\addcontentsline{toc}{subsection}{Processor Configuration Register}

The processor configuration register contains information about the current system and processor hardware setup. Some of its fields are read-only and set directly by the hardware, while others can be modified by processor-dependent code routines.

\begin{center}
\drawInstrFormatSixtyFour
  	{0, 8, 12, 13, 14, 15, 16}
  	{Cycles per millisecond, Flags,  PNum, RID, AId, MSize }
\end{center}


\begin{longtable}{@{}p{0.4\linewidth}p{0.5\linewidth}@{}}
    \caption{Processor Configuration Register Fields} \\
    \toprule
    \textbf{Field} & \textbf{Purpose} \\
    \midrule
    \endfirsthead
    \toprule
    \textbf{Field} & \textbf{Purpose} \\
    \midrule
    \endhead
    \midrule
    \multicolumn{2}{r}{\textit{Continued on next page}} \\
    \midrule
    \endfoot
    \bottomrule
    \endlastfoot
    \textbf{Memory Size (MSize)} & The available physical memory in 4 Gbyte increments. The minimum size is 4 Gbyte, encoded as a zero. The maximum size is 64 Gbyte encoded as a 15. \\
    \midrule
    \textbf{Architecture Id (AId) } & ... \\
    \midrule
    \textbf{Processor Id Id (RID) } & ... \\
    \midrule
    \textbf{Processor Core Num} & in a multiprocessor configuration, the unique processor number is assigned during system startup. \\
    \midrule
    \textbf{Flags (tbd) } & e.g default endian, etc. \\
    \midrule
    \textbf{Cycles per millisecond} & the cycle per millisecond specifies the number of processor cycles for a millisecond. \\
    \midrule
\end{longtable}


\subsection*{Recovery Counter}
\addcontentsline{toc}{subsection}{Processor Configuration Register}

When enabled, the recovery counter decrements on each instruction cycle. When the leftmost bit becomes a one and the R bit in the processor status word is set, a recovery trap is scheduled.

\subsection*{Shift Amount Register}
\addcontentsline{toc}{subsection}{Shift Amount Register}

The shift amount register ( \texttt{ SAR} ) is used by the extract, deposit and double shift instruction when a dynamic position is specified in the instruction. The valid number range is zero to 63.

\subsection*{Region Id Registers}
\addcontentsline{toc}{subsection}{Region Id Registers}

Twin-64 allows to record a set of up to 8 region IDs in the processor control registers \texttt{CR4} to \texttt{CR7} which are accessible to the currently executing process. For each access to a region that has the protection check bit set, one of the protection Id control registers must match the region Id. If not, a protection violation trap is raised. The rightmost bit of each of the eight RIDs is the write disable (W) bit. When the bit is set, that RID data cannot be used to grant write access. See also the chapter on addressing and access control.

\subsection*{Interrupt Vector Table Address}
\addcontentsline{toc}{subsection}{Interrupt Vector Table Address}

The Interruption Vector Table Address (IVA) control register, \texttt{CR8}, holds the base address of an array of service routines assigned to the different interruption classes. The lower 12 bits of the IVA are reserved and must be set to zero, meaning any address written to it must be aligned to a page boundary. The IVA is always located in  region zero. The array of interruption service routines is indexed by interruption numbers, as described in the chapter on interrupts.

\subsection*{External Interrupt Enable Mask}
\addcontentsline{toc}{subsection}{External Interrupt Enable Mask}

The External Interrupt Enable Mask (EIEM) is stored in \texttt{CR15)}. It is a 64-bit register, with each bit corresponding to one of the 64 external interrupts. A cleared bit in the EIEM masks the external interrupt associated with that position.

\subsection*{Interruption PSW}
\addcontentsline{toc}{subsection}{Interruption PSW}

Upon an interrupt, the address of the interrupted instruction and the processor status bits are saved to this control register.

\subsection*{Interruption Argument Registers}
\addcontentsline{toc}{subsection}{Argument Registers}

The Interruption Argument Registers (\texttt{CR11} and \texttt{CR12}) are used to pass an instruction and a virtual address to an interruption handler. The values of these registers for each interruption class are described in the chapter on interrupts.

\subsection*{Scratch Registers}
\addcontentsline{toc}{subsection}{External Interrupt Register}

Control registers \texttt{CR13}, \texttt{CR14}, and C\texttt{CR15} are scratch registers used by interrupt handlers or similar routines. \texttt{CR13} and \texttt{CR14} are privileged and can only be accessed from code running at a privileged level, whereas \texttt{CR15} is accessible from any privilege level. 

\section*{Program State Register}
\addcontentsline{toc}{section}{Processor State Register}

The processor state register contains the virtual address of the current instruction along with the processor status flags.

\begin{figure}[H]
\begin{center}
\begin{tikzpicture}[scale=0.7, transform shape]
        
    % \draw[help lines, gray!70, dashed] (0,0) grid(16,1.5);
  	
    \node[  tsRectangle, 
            minimum width=16cm,
            minimum height=1cm,  
            fill=white!50] (blocks) at (8,1) { };
    
    \def\fw{0.4}    
    
    \foreach \i in {1,...,12} {
    	\draw (\fw*\i,0.5) -- (\fw*\i,1.5);
	}       
                
  	\node at (\fw * 1 - 0.2, 1) {M};
	\node at (\fw * 2 - 0.2, 1) {E};
	\node at (\fw * 3 - 0.2, 1) {X};
	\node at (\fw * 4 - 0.2, 1) {R};
	\node at (\fw * 5 - 0.2, 1) {T};
	\node at (\fw * 6 - 0.2, 1) {-};
	\node at (\fw * 7 - 0.2, 1) {-};
	\node at (\fw * 8 - 0.2, 1) {-};
	\node at (\fw * 9 - 0.2, 1) {-};
	\node at (\fw * 10 - 0.2, 1) {C};
	\node at (\fw * 11 - 0.2, 1) {D};
	\node at (\fw * 12 - 0.2, 1) {I};
	
	\node at (\fw * 1 - 0.2,0.25) {\small{1}};
	\node at (\fw * 2 - 0.2,0.25) {\small{1}};
	\node at (\fw * 3 - 0.2,0.25) {\small{1}};
	\node at (\fw * 4 - 0.2,0.25) {\small{1}};
	\node at (\fw * 5 - 0.2,0.25) {\small{1}};
	\node at (\fw * 6 - 0.2,0.25) {\small{1}};
	\node at (\fw * 7 - 0.2,0.25) {\small{1}};
	\node at (\fw * 8 - 0.2,0.25) {\small{1}};
	\node at (\fw * 9 - 0.2,0.25) {\small{1}};
	\node at (\fw * 10 - 0.2,0.25) {\small{1}};
	\node at (\fw * 11 - 0.2,0.25) {\small{1}};
	\node at (\fw * 12 - 0.2,0.25) {\small{1}};
	
	\draw[-] (9,0.5) -- (9,1.5);
	
    \node at (7,0.25) {\small{20}};
   	\node at (12.5,0.25) {\small{32}};
   	
	\node at (6.5,1) {\small IA region};
	\node at (12.5,1) {\small IA Offset};
   	 	
\end{tikzpicture}
\end{center}
\caption{Program State Register}
\label{fig:Program State Register}
\end{figure}

\simpleNote{ Another option is to have a separate status word. Under evaluation. }

\subsection*{Instruction Address}
\addcontentsline{toc}{subsection}{Instruction Address Register}

The instruction address portion of the PSW registers contains the region id and the byte offset into the region for the currently executing instruction.

\subsection*{Processor Status}
\addcontentsline{toc}{subsection}{Processor Status}

\begin{longtable}{@{}p{0.25\linewidth}p{0.6\linewidth}@{}}
    \caption{Status Field Bits} \\
    \toprule
    \textbf{Bit} & \textbf{Purpose} \\
    \midrule
    \endfirsthead
    \toprule
    \textbf{Bit} & \textbf{Purpose} \\
    \midrule
    \endhead
    \midrule
    \multicolumn{2}{r}{\textit{Continued on next page}} \\
    \midrule
    \endfoot
    \bottomrule
    \endlastfoot
    \textbf{M} & High priority machine check. When set, high-priority machine checks are masked. This bit is set after an HPMC and cleared after all other interruptions. \\
    \midrule
    \textbf{E} & Little endian access enable. When set, memory access is in little endian format, else big endian format. Endian support is an optional feature. If not implemented, all access is in big endian format. \\
    \midrule
    \textbf{R} & Recovery counter enable. When set... \\
    \midrule
    \textbf{X} & When set, the processor runs in privileged mode. \\
    \midrule
    \textbf{T} & When set, taken branch traps are enabled. Any branch taken will result in a trap. \\
    \midrule
    \textbf{I} & When set, external interrupts are enabled. \\
    \midrule 
    \textbf{V} & Reserved. \\
    \midrule 
    \textbf{C} & Instruction memory break disable. The C-bit is cleared after the execution of each instruction, except for the \texttt{RFI} instruction which may set it. When set, instruction memory break traps are disabled. This bit allows a simple mechanism to trap on an instruction and then proceed past the trapping instruction. \\
    \midrule 
    \textbf{D} & Data memory break disable. The D-bit is cleared after the execution of each instruction, except for the \texttt{RFI} instruction which may set it. When set, data memory break traps are disabled. This bit allows a simple mechanism to trap on a data store and then proceed past the trapping instruction. \\
    \midrule
\end{longtable}

\section*{Data Types}
\addcontentsline{toc}{section}{Data Types}

The fundamental data type is a 64-bit signed integer, with the most significant bit on the left and the least significant bit on the right. All computations are performed on 64-bit signed integers. Memory accesses can be performed at the granularity of a double-word, word, halfword, or byte. Data loaded as a word, half--word, or byte is sign-extended to 64 bits. Store operations write the corresponding value to memory without modification.

\begin{figure}[ht]
\begin{center}
\begin{tikzpicture}[scale=0.7, transform shape]
        
  	% \draw[help lines, gray!70, dashed] (0,0) grid(16,5);
  	
  	\node[  tsRectangle, 
            minimum width=4cm,
            minimum height=0.75cm,
            fill=white!50] (blocks) at (14,4) { Byte };
                
    \node at (15.75,4.75) {0};
    \node at (12.25,4.75) {7};

    \node[  tsRectangle, 
            minimum width=8cm,
            minimum height=0.75cm,
            fill=white!50] (blocks) at (12,3) { Half };
                
    \node at (8.25,3.75) {15};
       
    \node[  tsRectangle, 
            minimum width=12cm,
            minimum height=0.75cm,
            fill=white!50] (blocks) at (10,2) { Word };
                
    \node at (4.25,2.75) {31};
    
    \node[  tsRectangle, 
            minimum width=16cm,
            minimum height=0.75cm,
            fill=white!50] (blocks) at (8,1) { Double };
            
    \node at (.25,1.75) {63};
                
      
\end{tikzpicture}
\end{center}
\caption{Data types}
\label{fig:data types}
\end{figure}

All memory addresses are expressed as bytes addresses. Address computation uses a 32-bit unsigned math, i.e. numbers wrap around in a 32-bit space and never cross region boundaries.

\section*{Byte Ordering and Alignment}
\addcontentsline{toc}{section}{Byte Ordering and Alignment}

Twin-64 is a big-endian machine. The following figure shows the memory access for load operations.

\begin{figure}[H]
\begin{center}
\begin{tikzpicture}[scale=0.9, transform shape]
        
  	% \draw[help lines, gray!70, dashed] (0,0) grid(15,7);
  	
  	\def\leftofs{2};
  	
  	\node at (1.5,6.75) {\large{Memory}};
  	\node at (11,6.75) {\large{Register}};

  	\node at (2+0.675, 6+0.25) {7};
	\node at (2+0.125, 6+0.25) {0};
  	\node at (1.5, 6 - 0.2) {0};
  	\node at (1.5, 0.2) {7};
  	
  	% Memory block
  	\node[tsRectangle, minimum width=0.75cm, minimum height=8*0.75cm, anchor=west] 
  	    at (\leftofs, 3) {};

    % Horizontal separators
    \foreach \y in {1,...,7} {
        \draw (\leftofs,\y*0.75) -- (\leftofs+0.75,\y*0.75);
    }

    % Memory labels a–h    
     \foreach \i/\lbl in {1/a,2/b,3/c,4/d,5/e,6/f,7/g,8/h} {
        \node at (\leftofs+0.375,\i*0.75-0.375) {\lbl};
    }

	% Byte
 	\node[  tsRectangle, 
            minimum width=8cm,
            minimum height=0.75cm,  
            anchor=west] at (7,5.5) { };
            
    \node[anchor=west] at ( 4, 5.5 ) { LD.B R3,7(R4) };        
    
    \foreach \i in {8,...,14} {
    	\draw[-] (\i, 5.5+0.375 ) -- (\i, 5.5 - 0.375);
    } 
    
    \foreach \i/\lbl in {7.5/0,8.5/0,9.5/0,10.5/0,11.5/0,12.5/0,13.5/0,14.5/a} {
        \node at (\i, 5.5) {\lbl};
    }
    
    % Half
    \node[  tsRectangle, 
            minimum width=8cm,
            minimum height=0.75cm,  
            anchor=west] at (7, 3.875) { };
            
    \node[anchor=west] at ( 4, 3.875 ) { LD.H R3,2(R4) };
    
    \foreach \i in {8,...,14} {
    	\draw[-] (\i, 3.875 + 0.375) -- (\i, 3.875 - 0.375);
    } 
    
    \foreach \i/\lbl in {7.5/0,8.5/0,9.5/0,10.5/0,11.5/0,12.5/0,13.5/f,14.5/e} {
        \node at (\i, 3.875) {\lbl};
    }
        
    % Word
    \node[  tsRectangle, 
            minimum width=8cm,
            minimum height=0.75cm,  
            anchor=west] at (7,2.25) { };
            
    \node[anchor=west] at ( 4, 2.25 ) { LD.W R3,4(R4) };
    
    \foreach \i in {8,...,14} {
    	\draw[-] (\i, 2.25 + 0.375) -- (\i, 2.25 - 0.375);
    } 
    
    \foreach \i/\lbl in {7.5/0,8.5/0,9.5/0,10.5/0,11.5/d,12.5/c,13.5/b,14.5/a} {
        \node at (\i, 2.25) {\lbl};
    }

 	% Double
    \node[  tsRectangle, 
            minimum width=8cm,
            minimum height=0.75cm,  
            anchor=west] at (7, 0.5) { };
            
    \node[anchor=west] at ( 4, 0.5 ) { LD.D R3,0(R4) }; 
    
    \foreach \i in {8,...,14} {
    	\draw[-] (\i, 1 - 0.125) -- (\i, 0 + 0.125);
    } 
    
    \foreach \i/\lbl in {7.5/h,8.5/g,9.5/f,10.5/e,11.5/d,12.5/c,13.5/b,14.5/a} {
        \node at (\i, 0.5) {\lbl};
    }
    
\end{tikzpicture}
\end{center}
\caption{Big endian mode}
\label{fig:Big endian mode}
\end{figure}

The processor status register reserves a bit position, "E", for future support of mixed big and little endian memory access. Currently this bit is always set to zero.

Twin-64 enforces natural alignment for memory accesses. Byte, half-word, word, and double-word data types must be aligned to 1, 2, 4, and 8-byte boundaries respectively. If an access does not meet the required alignment, a data-alignment trap is raised.

\section*{Translation Lookaside Buffer}
\addcontentsline{toc}{section}{Translation Lookaside Buffer}

Virtual address translations are stored in a translation lookaside buffer (TLB). The TLB is essentially a cache of page table entries representing the virtual pages currently mapped to physical addresses. Depending on the hardware implementation, there may be a single combined TLB or separate instruction and data TLBs. Each TLB entry stores the virtual page number (VPN) along with its attributes and the corresponding physical page number (PPN). The current architecture supports a 36-bit physical address space, allowing a maximum of 64 GB of memory.

\begin{figure}[H]
\begin{center}
\begin{tikzpicture}[scale=0.9, transform shape]
        
  	% \draw[help lines, gray!70, dashed] (0,0) grid(16,5);
  	
    \node[  tsRectangle, 
            minimum width=10cm,
            minimum height=0.75cm,  
            fill=white,
            anchor=west] at (0,4.5) { };
    \node[anchor=west] at (0, 4.5) {Virtual page number};
    \node[anchor=west] at (11,4.5) {38bit};
            
  	\node[  tsRectangle, 
            minimum width=6cm,
            minimum height=0.75cm,  
            fill=white,
            anchor=west] at (0,3.5) { };
    \node[anchor=west] at (0, 3.5) {Physical page number};   
    \node[anchor=west] at (11,3.5) {24bit};
     
    \node[  tsRectangle, 
            minimum width=2cm,
            minimum height=0.75cm,  
            fill=white,
            anchor=west] at (0,2.5) { };
    \node[anchor=west] at (0, 2.5) {AccRights}; 
    \node[anchor=west] at (11,2.5) {4bit};
    
    \node[  tsRectangle, 
            minimum width=2cm,
            minimum height=0.75cm,  
            fill=white,
            anchor=west] at (0,1.5) { };
    \node[anchor=west] at (0, 1.5) {PSize}; 
    \node[anchor=west] at (11,1.5) {4bit};
    
    \def\fw{0.4} 
    
    \node[  tsRectangle, 
            minimum width=6*\fw cm,
            minimum height=0.75cm,  
            fill=white,
            anchor=west] at (0,0.5) { };
           
    \foreach \i/\lbl in {0/V,1/M,2/L,3/U,4/P,5/B } {
    	\draw (\fw*\i,0.125) -- (\fw*\i,1 - 0.125);
    	\node at (\fw*\i + 0.2, 0.5 ) {\lbl};
	}         
            
    \node[anchor=west] at (11,0.5) {6bit};
    
\end{tikzpicture}
\end{center}
\caption{Tlb info}
\label{fig: Tlb info}
\end{figure}

To reduce TLB processing overhead, the TLB is capable of storing virtual address ranges. There is support for a larger page sizes. The supported sizes are 4 Kbyte, 64 Kbyte, 1 MByte and 16Mbyte. 

\begin{longtable}{@{}p{0.25\linewidth}p{0.6\linewidth}@{}}
    \caption{TLB Fields} \\
    \toprule
    \textbf{Field} & \textbf{Purpose} \\
    \midrule
    \endfirsthead
    \toprule
    \textbf{Field} & \textbf{Purpose} \\
    \midrule
    \endhead
    \midrule
    \multicolumn{2}{r}{\textit{Continued on next page}} \\
    \midrule
    \endfoot
    \bottomrule
    \endlastfoot
    \textbf{VPN} & The virtual page number. \\
    \midrule
    \textbf{PPN} & The physical page number. \\
    \midrule
    \textbf{Access Rights} & Encodes the access type (read, write, and execute) and the required privilege level for the operation. \\
    \midrule
    \textbf{Page Size} & 4-bit field encoding the supported page size: 
    		\[ \texttt{( 1U << ( pSize * 4 )) * T64\_PAGE\_SIZE} \] 
    The pSize number 4 to 15 are reserved and result in an undefined operation. A page size is not allowed to cross region boundaries. The virtual address must be aligned with selected the page size.\\
    \midrule
    \textbf{V} & Indicates if the entry is valid. \\
    \midrule
    \textbf{M} & When set, the page has been modified. This only applies to data TLBs or unified TLBs. \\
    \midrule
    \textbf{L} & When set, the TLB entry is locked and will not be selected during replacement. \\
    \midrule
    \textbf{U} & When set, the page represented by the TLB entry is not cached. \\
    \midrule
    \textbf{B} & When set, a branch instruction executed from this page will cause a branch-taken trap. This inly applies for instruction TLBs or unified TLBs.\\
\end{longtable}

The first regions, region 0 to 15, are special in that they bypass the TLB entirely. A virtual address in the range 0x00\_0000\_0000 to 0x03\_FFFF\_FFFF directly maps to the physical address of the same range. The address range represents physical memory and can only be accessed in privileged mode. 

\section*{Caches}
\addcontentsline{toc}{section}{Caches}

Twin-64 is a cache-visible architecture that supports both unified and separate instruction and data caches. While the hardware directly manages cache line insertion and replacement, software has explicit control over flushing and purging cache line entries. Twin-64 caches are virtually indexed and physically tagged. To accelerate memory access, the cache uses the virtual address to index into the arrays in parallel with the TLB lookup. A cache hit occurs when the physical address tag in the cache matches the physical page address from the TLB.

The architecture does not mandate specific cache line sizes, cache organization, or hierarchy models. Instead, it defines instructions for cache operations, such as deletion or flushing of cache lines.

Each TLB entry indicates whether the corresponding data should be cached. For regions 0 to 15, where the TLB is bypassed, the hardware determines the caching behavior. All memory address ranges as well as the I/O address range are not cached.

\section*{Instruction Set}
\addcontentsline{toc}{section}{Instruction Set}

Twin-64 features a simple and efficient instruction set. All instructions are of fixed length, specifically a 32-bit word. The total number of instructions is relatively small, but many include execution options that make them highly versatile. Great care has been taken to ensure these options do not significantly complicate the pipeline or lengthen the data path.  

Unlike CISC-style processors that support a wide variety of addressing modes, Twin-64 provides just two virtual memory addressing modes, both based on a simple base–offset calculation. No indirection or mode requiring a memory operand to compute an address is part of the architecture.

The instruction set is organized into five groups:  

\begin{itemize}
  \item \textbf{Computational instructions:} Arithmetic, boolean, and bit-manipulation operations such as extract and deposit. Twin-64 supports only 64-bit signed arithmetic. Any numeric overflow results in a trap.
 
  \item \textbf{Immediate instructions:} Instructions for constructing immediate values up to 64 bits. Several instructions embed immediate fields within the instruction word, but a full 64-bit constant requires a sequence of two or three instructions. There are also immediate instructions that support 32-bit address arithmetic, where calculations wrap around within the 32-bit offset field and never overflow into the region identifier portion.
 
  \item \textbf{Memory reference instructions:} Load and store operations for both virtual and physical memory. These transfer data between memory and general registers in sizes of double--word, word, halfword, or byte. The design resembles a load/store architecture, but unlike strict load/store machines, many Twin-64 instructions allow memory operands directly. No instruction, however, both reads and writes data memory in the same operation.
  
  \item \textbf{Control flow instructions:} Both unconditional and conditional branches are supported. Unconditional branches may optionally save the return address in a general-purpose register. Conditional branches compare two operands and transfer control if the condition is satisfied, eliminating the need for a separate compare instruction.
    
  \item \textbf{System control instructions:} Instructions for managing processor resources such as the TLB, cache, and control registers. Control registers can be accessed directly; the TLB can be updated with insert and remove operations; and the cache can be flushed or purged. Additional instructions generate traps or provide diagnostic functions for examining processor state. Most system instructions require execution in privileged mode.
\end{itemize}

The book section on the instruction set will describe each instruction group in detail, with formats, descriptions, and examples.
