\chapter*{Simulator Configuration}
\addcontentsline{toc}{chapter}{Simulator Configuration}

The simulator can be configured using a set of global options that control its
startup behavior and runtime environment.

Configuration information can be supplied through multiple mechanisms:
\begin{itemize}
    \item Program command-line parameters
    \item Environment variables
    \item Simulator commands executed during initialization
\end{itemize}

These mechanisms work together to define the initial simulator state before
interactive use begins.

%--------------------------------------------------------------------------------
\section*{Program Parameters}
\addcontentsline{toc}{section}{Program Parameters}

Program parameters are specified on the simulator command line and control
basic startup behavior, such as output verbosity, configuration file selection,
and version or help display.

The following table lists the supported command-line parameters.

\begin{table}[h]
\centering
\begin{tabularx}{\linewidth}{|l|l|X|}
\hline
\textbf{Name} & \textbf{Type} & \textbf{Usage} \\ \hline
\texttt{help} &
No argument &
Displays a summary of available command-line options and exits. \\ \hline

\texttt{version} &
No argument &
Displays the simulator version information and exits. \\ \hline

\texttt{verbose} &
No argument &
Enables verbose output during simulator initialization and execution. \\ \hline

\texttt{config} &
Required argument &
Specifies the path to a configuration file containing simulator commands to be
executed at startup. \\ \hline

\texttt{log} &
Required argument &
Specifies the path to the log file used for runtime logging. \\ \hline
\end{tabularx}
\caption{Simulator command-line options}
\end{table}

%--------------------------------------------------------------------------------
\section*{Configuration Files}
\addcontentsline{toc}{section}{Configuration Files}

In addition to command-line parameters, the simulator can be configured using
a configuration file specified with the \texttt{--config} program option.

A configuration file contains a sequence of simulator commands that are
executed automatically after the initial program setup is complete. This
mechanism allows complex simulator setups to be reproduced reliably and
consistently.

Configuration files are typically used to:
\begin{itemize}
    \item Set environment variables
    \item Create and initialize simulator windows
    \item Instantiate Twin-64 system modules
    \item Establish an initial debugging or inspection layout
\end{itemize}

After all commands in the configuration file have been executed, control is
passed to the interactive command interface.

%--------------------------------------------------------------------------------
\section*{Configuration Example}
\addcontentsline{toc}{section}{Configuration Example}

to do ...