%--------------------------------------------------------------------------------
%
%
%--------------------------------------------------------------------------------
\chapter*{Addressing and Access Control}
\addcontentsline{toc}{chapter}{Addressing and Access Control}



\section*{Address Translation}
\addcontentsline{toc}{section}{Address Translation}

The Twin-64 CPU emits always a virtual address. Each memory access needs to be translated into a physical address.

virtual address for the first four segments directly map to the physical address space.

Twin-64 implements a protection model along two dimensions. 

\section*{Access Control}
\addcontentsline{toc}{section}{Access Control}

The first dimension is the \textbf{access control} and \textbf{privilege level} check. Each page is associated with a page type. There are read-only, read-write, execute and gateway pages. Each memory access is checked to be compatible with the allowed type of access set in the page descriptor. There are two privilege levels, user and supervisor mode. Access type and privilege level form the access control information field, which is checked for each instruction and data memory access. For read access the privilege level us be least as high as the PL1 field, for write access the privilege level must be as least as high as PL2. An exception are the execute and gateway pages, where a write access is only allowed for privilege code. For execute access the privilege level must be at least as high as PL1 and not higher than PL2. If the instruction privilege level is not within this bounds, a privilege violation trap is raised. If the page type does not match the instruction access type an access violation trap is raised. In both cases the instruction is aborted and a memory protection trap is raised.

\section*{Access Protection}
\addcontentsline{toc}{section}{Access Protection}

The second dimension of protection is a \textbf{protection ID}. Twon-64 allows to record a set of segment Id values in the processor control registers. For each access to a segment that has the protection check bit set, one of the protection Id control registers must match the segment Id. If not, a protection violation trap is raised.

Protection Id checking is typically used to form a grouping of access. A good example is the stack data segment, which is accessible at user privilege level. Since the CPU features a global virtual memory address space, every task could access a data stack of another task. With protection Id checking enabled and the segment Id loaded in one of the control registers, only the corresponding user task is allowed to access the stack data segment with a matching protection Id. 

