%--------------------------------------------------------------------------------
%
%
%--------------------------------------------------------------------------------
\chapter*{Instruction Set Descriptions}
\addcontentsline{toc}{chapter}{Instruction Set Descriptions}

This chapter provides a detailed description of the Twin-64 instruction set. Each instruction entry includes:

\begin{itemize}
	\setlength{\itemsep}{0.2em}
    \setlength{\topsep}{0pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}
    \item The instruction word layout.
    \item A concise description of the instruction.
    \item A high-level pseudo-code representation of its operation.
\end{itemize}

\section*{Instruction Operations}
\addcontentsline{toc}{section}{Instruction Operations}

Instruction operations are expressed using C-style pseudo code. The conventions and helper routines used in these descriptions are defined in the appendix. Registers are referenced by class and index:

\begin{itemize}
    \setlength{\itemsep}{0.2em}
    \setlength{\topsep}{0pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}
    \item \texttt{GR[5]} — general register 5
    \item \texttt{CR[1]} — control register 1
\end{itemize}

Some registers have dedicated names; for example, the shift-amount control register is \texttt{SAR}.

\section*{Instruction Options}
\addcontentsline{toc}{section}{Instruction Options}

Instruction options are specified in assembler syntax by appending a dot to the instruction mnemonic, followed by one or more option characters. Each character represents a defined option. For example:

\begin{itemize}
    \setlength{\itemsep}{0.2em}
    \setlength{\topsep}{0pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}
    \item \texttt{AND.N ...} — the \texttt{AND} instruction with the negate-result option.
    \item \texttt{LD.B ... } - the texttt{LD} instruction with loading a byte option.    
\end{itemize}

The order of option characters is insignificant. All valid options for an instruction are listed in the corresponding option set.

\section*{Instruction Subfields}
\addcontentsline{toc}{section}{Instruction Subfields}

Individual subfields of an instruction or register are accessed using a dot followed by the field name enclosed in square brackets. For example:

\begin{itemize}
    \setlength{\itemsep}{0.2em}
    \setlength{\topsep}{0pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}
    \item The interrupt-enable bit in the program state register: \texttt{PSR.[I]}.
\end{itemize}

Fields marked as \emph{reserved} are intended for future extensions and must always be set to zero. Any non-zero value in a reserved field results in undefined behavior.

\section*{Constant Expressions}
\addcontentsline{toc}{section}{Constant Expressions}

Constant expressions are 64-bit values. When encoding a numeric field that cannot be represented directly, a sequence of instructions may be required to construct the value in the destination word. To support this, constant values may be preceded by a field qualifier that selects a portion of the 64-bit value.

The available field qualifiers are \texttt{U\%}, \texttt{M\%}, \texttt{L\%}, and \texttt{R\%}. Each qualifier, when applied to a 64-bit numeric value, extracts the corresponding field and right-shifts it to form the resulting expression.

The following figure illustrates the individual fields selected by each qualifier:

\begin{center}
    \drawInstrFormatSixtyFour
       {0, 3, 8, 13, 16}
        {U, M, L, R}
\end{center}

Examples of constant expression usage can be found in the descriptions of the \texttt{ADDIL}, \texttt{LDIL}, and \texttt{MFIA} instructions

%--------------------------------------------------------------------------------
%
%
%--------------------------------------------------------------------------------
\input {chapter-instruction-set-descriptions/Twin-64-Opcodes}
\input {chapter-instruction-set-descriptions/section-instruction-ABR}
\input {chapter-instruction-set-descriptions/section-instruction-ADD}
\input {chapter-instruction-set-descriptions/section-instruction-ADDIL}
\input {chapter-instruction-set-descriptions/section-instruction-AND}
\input {chapter-instruction-set-descriptions/section-instruction-B}
\input {chapter-instruction-set-descriptions/section-instruction-BB}
\input {chapter-instruction-set-descriptions/section-instruction-BE}
\input {chapter-instruction-set-descriptions/section-instruction-BR}
\input {chapter-instruction-set-descriptions/section-instruction-BV}
\input {chapter-instruction-set-descriptions/section-instruction-CBR}
\input {chapter-instruction-set-descriptions/section-instruction-CMP}
\input {chapter-instruction-set-descriptions/section-instruction-DEP}
\input {chapter-instruction-set-descriptions/section-instruction-DIAG}
\input {chapter-instruction-set-descriptions/section-instruction-DSR}
\input {chapter-instruction-set-descriptions/section-instruction-EXTR}
\input {chapter-instruction-set-descriptions/section-instruction-FCA}
\input {chapter-instruction-set-descriptions/section-instruction-ITLB}
\input {chapter-instruction-set-descriptions/section-instruction-LD}
\input {chapter-instruction-set-descriptions/section-instruction-LDIL}
\input {chapter-instruction-set-descriptions/section-instruction-LDO}
\input {chapter-instruction-set-descriptions/section-instruction-LDR}
\input {chapter-instruction-set-descriptions/section-instruction-LPA}
\input {chapter-instruction-set-descriptions/section-instruction-MBR}
\input {chapter-instruction-set-descriptions/section-instruction-MFIA}
\input {chapter-instruction-set-descriptions/section-instruction-MFCR}
\input {chapter-instruction-set-descriptions/section-instruction-MTCR}
\input {chapter-instruction-set-descriptions/section-instruction-NOP}
\input {chapter-instruction-set-descriptions/section-instruction-OR}
\input {chapter-instruction-set-descriptions/section-instruction-PCA}
\input {chapter-instruction-set-descriptions/section-instruction-PRB}
\input {chapter-instruction-set-descriptions/section-instruction-PTLB}
\input {chapter-instruction-set-descriptions/section-instruction-RFI}
\input {chapter-instruction-set-descriptions/section-instruction-RSM}
\input {chapter-instruction-set-descriptions/section-instruction-SHLXA}
\input {chapter-instruction-set-descriptions/section-instruction-SHRXA}
\input {chapter-instruction-set-descriptions/section-instruction-SSM}
\input {chapter-instruction-set-descriptions/section-instruction-ST}
\input {chapter-instruction-set-descriptions/section-instruction-STC}
\input {chapter-instruction-set-descriptions/section-instruction-SUB}
\input {chapter-instruction-set-descriptions/section-instruction-TRAP}
\input {chapter-instruction-set-descriptions/section-instruction-XOR}
