%--------------------------------------------------------------------------------
%
%
%--------------------------------------------------------------------------------
\pagebreak
\section*{ST — Store Register}
\addcontentsline{toc}{section}{ST — Store Register}

\begin{iPageDescEntry}{Syntax}
    \texttt{ST[.D/W/H/B] RegR, SignedImmed13( RegB )} \\
    \texttt{ST[.D/W/H/B] RegR, RegX( RegB )}
\end{iPageDescEntry}

\begin{iPageDescEntry}{Format}
    The \texttt{ST} instruction uses the indexed and register-indexed instruction formats. \\

    \begin{flushleft}
    \drawInstrFormat
        {0,1,3,5,5.5,6,6.5,8.5,9.5,16}
        {\OpCodeGrpMEM, \OpCodeST, RegR, 0, M, 0, RegB, dw, S-IMM-13}
    \end{flushleft}

    \begin{flushleft}
    \drawInstrFormat
        {0,1,3,5,5.5,6,6.5,8.5,9.5,11.5,16}
        {\OpCodeGrpMEM, \OpCodeST, RegR, 0, M, 1, RegB, dw, RegX, 0}
    \end{flushleft}
\end{iPageDescEntry}

\begin{iPageDescEntry}{Description}
    The \texttt{ST} instruction stores a value from a general-purpose register into memory. 
    
    In the immediate-offset format, the signed immediate offset is shifted right by the \texttt{dw} field of the instruction and added to \texttt{RegB} to form the effective address.  
    In the register-indexed format, \texttt{RegX} is added to \texttt{RegB} to form the effective address.
    
    Optionally, the computed effective address may also be written back to the base register \texttt{RegB}. When this option is enabled, the sign of the offset determines the update behavior:  
    – A negative offset causes the base register itself to be used as the fetch address.  
    – A positive offset causes the sum of the base register and offset to be used.  
    In both cases, the base register is updated.
    
    The store address must be aligned to the data width specified. The value stored to memory is stored  as is in the data width specified.
\end{iPageDescEntry}

\begin{iPageDescEntry}{Operation}
\begin{lstlisting}[style=iPageOpStyle]
memStore( memOperand( addOfs32( RegB, ofs )), RegR );
memStore( memOperand( addOfs32( RegB, RegX )), RegR );
\end{lstlisting}
\end{iPageDescEntry}

\begin{iPageDescEntry}{Exceptions}
    - Data Alignment Trap \\
    - Memory Access Violation Trap \\
    - Memory Protection Violation Trap \\
    - Data TLB Miss Trap
\end{iPageDescEntry}

\begin{iPageDescEntry}{Notes}
    The jury is still out if the M bit is not too expensive in hardware. We would need to read three registers. One for the value, and up to two for the index calculation.
\end{iPageDescEntry}


