%--------------------------------------------------------------------------------
%
%
%--------------------------------------------------------------------------------
\pagebreak
\section*{STC — Store Register Conditional}
\addcontentsline{toc}{section}{STC — Store Register Conditional}

\begin{iPageDescEntry}{Syntax}
  \texttt{STC RegR, SignedImmed13(RegB)}
\end{iPageDescEntry}

\begin{iPageDescEntry}{Format}
    The \texttt{STC} instruction uses the indexed instruction format. \\

    \begin{flushleft}
    \drawInstrFormat
        {0, 1, 3, 5, 6.5, 8.5, 9.5, 16}
        {\OpCodeGrpMEM, \OpCodeSTC, RegR, 0, RegB, 3, S-IMM-13}
    \end{flushleft}
\end{iPageDescEntry}

\begin{iPageDescEntry}{Description}
    The \texttt{STC} instruction stores a value from a general-purpose register to memory
    only if the memory address is still reserved by a prior \texttt{LDR} instruction.
    - If the reservation is valid, the store succeeds and clears the reservation.
    - If the reservation has been lost (e.g., another processor wrote to the address), the store fails and the memory is unchanged.
    
    May cause a memory reference trap if the address is invalid.
\end{iPageDescEntry}

\begin{iPageDescEntry}{Operation}
    \texttt{if (isAddressReserved(memOperand())) \{ } \\
    \quad \texttt{memStore(memOperand(), RegR);} \\
    \quad \texttt{clearReservation(memOperand());} \\
    \texttt{\} else \{ } \\
    \quad \texttt{storeFailedFlag <- 1;} \\
    \texttt{\}}
\end{iPageDescEntry}

\begin{iPageDescEntry}{Exceptions}
    - Data Alignment Trap \\
    - Memory Access Violation Trap \\
    - Memory Protection Violation Trap \\
    - Data TLB Miss Trap
\end{iPageDescEntry}

\begin{iPageDescEntry}{Notes}
    - This instruction is the second part of a *load-reserved / store-conditional* pair for atomic memory updates. \\
    - Success or failure of the store is typically tested via a status flag. \\
    - Commonly used for implementing spinlocks, mutexes, and atomic counters in multiprocessor systems.
\end{iPageDescEntry}



