%--------------------------------------------------------------------------------
\pagebreak
\section*{BE - Branch External}
\addcontentsline{toc}{section}{BE - Branch External}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Syntax}
\texttt{BE [ ofs ] ( RegB ) [, RegR ]}
\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Format}
The \texttt{BE} instruction uses the signed immediate S15 format. \\

\begin{flushleft}
\drawInstrFormat
    {0,1,3,5,6.5,8.5,16}
    {\OpCodeGrpBR,\OpCodeBV, RegR, 0, RegB, ofs}
\end{flushleft}
\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Description}
The \texttt{BE} instruction performs an unconditional branch to another segment. \texttt{RegB} contains a target segment and offset base to which the sign extended offset shifted by two is added. The return link is stored in \texttt{RegR}.

Because \texttt{BE} is a segment base relative branch, branching to a page with a different privilege level is possible. Branching from a lower privilege level to a higher one triggers an instruction protection trap. Branching from a higher privilege to a lower privilege level updates the privilege level in the processor status register. If no change is required, the privilege level remains unchanged.  


\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescOpEntry}{Operation}
\begin{lstlisting}[style=iPageOpStyle]
RegR <- addOfs32( PSW.[IA], 4 );
PSW.[IA] <- addOfs32( RegB, RegX << 2 );
\end{lstlisting}
\end{iPageDescOpEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Exceptions}
Taken Branch Trap\\
Instruction TLB Miss Trap
\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Notes}
None.
\end{iPageDescEntry}

