%--------------------------------------------------------------------------------
%
%
%--------------------------------------------------------------------------------
\pagebreak
\section*{CBR - Compare and Branch}
\addcontentsline{toc}{section}{CBR - Compare and Branch}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Syntax}

	\texttt{CBR.<cond> RegR, RegB, target}
    
\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Format}

The CBR instruction uses the signed immediate S15 format:
	
\begin{flushleft}
\drawInstrFormat
  	{0,1,3,5,6.5,8.5,16}
  	{\OpCodeGrpBR,\OpCodeCBR,Reg R, cond, Reg B, ofs }
\end{flushleft}
\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Description}

The \texttt{CBR} instruction compares the contents of \texttt{RegB} with \texttt{RegR} using the condition specified in the \texttt{cond} field. If the condition is satisfied, control is transferred to the instruction address computed by adding the sign-extended and word-aligned offset to the current instruction address. Otherwise, execution continues with the next sequential instruction. The comparison operation uses the same signed 64-bit comparator as the \texttt{CMP} instruction, and neither source register is modified.
\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Conditions}

The \texttt{cond} field encodes the branch condition as follows:

\begin{center}
\begin{tabular}{|c|c|l|}
\hline
\textbf{Value}  & \textbf{Mnemonic} & \textbf{Branch if} \\ \hline
0 & \texttt{EQ} & \texttt{RegR == RegB} \\ \hline
1 & \texttt{LT} & \texttt{RegR <  RegB} \\ \hline
2 & \texttt{GT} & \texttt{RegR >  RegB} \\ \hline
3 & reserved    & \\ \hline
4 & \texttt{NE} & \texttt{RegR != RegB} \\ \hline
5 & \texttt{GE} & \texttt{RegR >= RegB} \\ \hline
6 & \texttt{LE} & \texttt{RegR <=  RegB} \\ \hline
7 & reserved    & \\ \hline
\end{tabular}
\end{center}

\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescOpEntry}{Operation}
\begin{lstlisting}[style=iPageOpStyle]
if ( cond( RegR, RegB )) 
   PSW.[IA-OFS] <- addOfs32( PSW.[IA-OFS], ( ofs << 2 ));
else                     
   PSW.[IA-OFS] <- addOfs32( PSW.[IA-OFS] + 4 );
\end{lstlisting}
\end{iPageDescOpEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Exceptions}
	Taken Branch Tap.\\
	Instruction TLB Miss.
\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Notes}{
The \texttt{CBR} instruction extends the \texttt{CMP} functionality by integrating a conditional branch. With a mode field of three bits available for the \texttt{cond} field, all six standard relational conditions (\texttt{EQ}, \texttt{NE}, \texttt{LT}, \texttt{LE}, \texttt{GT}, \texttt{GE}) are directly encoded, avoiding the need for operand swapping or result inversion.

See also the instruction notes on comparison and branch Condition encoding in the appendix.
}
\end{iPageDescEntry}
