%--------------------------------------------------------------------------------
%
%
%--------------------------------------------------------------------------------
\pagebreak
\section*{IxTLB — Insert TLB Entry}
\addcontentsline{toc}{section}{IxTLB — Insert TLB Entry}

\begin{iPageDescEntry}{Syntax}
    \texttt{IITLB RegR, RegB, RegA} \\
    \texttt{IDTLB RegR, RegB, RegA}
\end{iPageDescEntry}

\begin{iPageDescEntry}{Format}
    The \texttt{IITLB} instruction uses the following instruction format. \\

    \begin{flushleft}
    \drawInstrFormat
        {0,1,3,5,6.5,8.5,9.5,11.5,16}
        {\OpCodeGrpSYS, \OpCodePTLB, RegR, 0, RegB, 0, RegA, 0}
    \end{flushleft}
    
    The \texttt{IDTLB} instruction uses the following instruction format. \\

    \begin{flushleft}
    \drawInstrFormat
        {0, 1,3,5,6.5,8.5,9.5,11.5,16}
        {\OpCodeGrpSYS, \OpCodePTLB, RegR, 1, RegB, 0, RegA, 0}
    \end{flushleft}

\end{iPageDescEntry}

\begin{iPageDescEntry}{Description}
    The \texttt{IxTLB} instruction inserts a new entry into the translation lookaside buffer (xTLB). The virtual address is stored in \texttt{RegB} and additional control information is provided by \texttt{RegA}. This is a privileged instruction.
\end{iPageDescEntry}

\begin{iPageDescEntry}{TLB Arguments}
	\texttt{RegB}  and \texttt{RegA} contain the arguments passed to the TLB subsystem. \texttt{RegB}  contains the virtual address. The address is always o a page boundary and in the case of address ranges marks the starting page.
	
	\begin{flushleft}
	\drawInstrFormatSixtyFour
        {0, 3, 13,16}
        {0, VPN, 0}
    \end{flushleft}
    
    \texttt{RegA} contains the flags, access rights, page size and the physical address of the address range. The appendix contains the definitions of the TLB info word.
    
    \begin{flushleft}
	\drawInstrFormatSixtyFour
        {0,0.25,0.5, 0.75, 1, 1.25, 1.5, 1.75, 2, 3, 4, 7, 13, 16}
        {0, L,U, M, -, -, -, -, Acc, Size, 0, PPN, 0}
    \end{flushleft}
    
    The L bit will lock the entry in the TLB. When looking for a free entry, these entries are skipped. If all entries are locked entry 0, TLB entry zero is selected for replacement. The U bit will mark the address range described by the TLB entry as uncached. The M bit is set when a store type instruction access the translation.
    
    The TLB arguments layout shows a page size of 4Kb. Depending on the page size specified in the size field, the VPN and PPN fields are shorter and the offset is larger. The sum of both fields is however always 52 bits for virtual pages, and 36 bits for physical pages.
\end{iPageDescEntry}

\begin{iPageDescOpEntry}{Operation}
\begin{lstlisting}[style=iPageOpStyle]
RegR <- insertIntoTlb( RegB, RegA );
\end{lstlisting}
\end{iPageDescOpEntry}

\begin{iPageDescEntry}{Exceptions}
    Privilege Violation Trap.
\end{iPageDescEntry}

\begin{iPageDescEntry}{Notes}
    None.  
\end{iPageDescEntry}
