%--------------------------------------------------------------------------------
%
%
%--------------------------------------------------------------------------------
\pagebreak
\section*{LDIL — Load Immediate Left}
\addcontentsline{toc}{section}{LDIL — Load Immediate Left}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Syntax}
\texttt{LDIL[.L/.M/.U] RegR, val}
\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Format}
The \texttt{LDIL} instruction uses the following instruction format: \\

\begin{flushleft}
\drawInstrFormat
    {0,1,3,5,6,16}
    {\OpCodeGrpALU,\OpCodeLDIL, RegR, mode, val}
\end{flushleft}
\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Description}
The \texttt{LDIL} instruction deposits an immediate value into general register \texttt{RegR}.  
The optional suffix determines how the immediate value is positioned within the destination register. If no suffix is specified, the default is \texttt{.L} mode (mode value = 1).

\begin{itemize}
    \setlength{\itemsep}{0.5em}
    \setlength{\parskip}{0pt}
    \item \texttt{.L} (mode = 1) — Deposits 20 bits of \texttt{val} into bits [31..12] of \texttt{RegR}. RegR is cleared before the operation. 
    \item \texttt{.M} (mode = 2) — Deposits 20 bits of \texttt{val} into bits [51..32] of \texttt{RegR}. RegR is not cleared before the operation.
    \item \texttt{.U} (mode = 3) — Deposits 12 bits of \texttt{val} into bits [63..52] of \texttt{RegR}. RegR is not cleared before the operation.
\end{itemize}

A mode value of zero is reserved for the \texttt{ADDIL} instruction and must not be used here.
\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescOpEntry}{Operation}
\begin{lstlisting}[style=iPageOpStyle]
RegR <- ((val & 0xFFFFF) << 12 );                 // .L mode (1)
RegR <- deposit( RegR, (val & 0xFFFFF), 32, 20 ); // .M mode (2)
RegR <- deposit( RegR, (val & 0xFFF), 52, 12 );   // .U mode (3)
\end{lstlisting}
\end{iPageDescOpEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Exceptions}
None.
\end{iPageDescEntry}

%--------------------------------------------------------------------------------
\begin{iPageDescEntry}{Notes}
The \texttt{LDIL} instruction is used to construct large values in a register.  
The \texttt{.L} variant clears the destination register before depositing the immediate value.  It is the default option for the \texttt{LDIL} instruction.
This form is intended for building larger offsets for indexed addressing modes, where an addressing instruction adds its own 12-bit offset in the low bits. Together, these operations allow the construction of any 32-bit offset.

In contrast, the \texttt{.M} and \texttt{.U} variants deposit their immediate values without clearing the register, preserving the remaining bits. Using a combination of these modes, a full 64-bit value can be assembled in two to four instructions.

For example constructing a 52-bit value, the instruction sequence \\

\begin{lstlisting}[style=iPageOpStyle]
LDIL   R1, L%0xA_BCDE_1234_5678  ; load    0x1234_5000 into R1
ADDIL  R1, R%0xA_BCDE_1234_5678  ; add     0x678       into R1
LDIL.M R1, M%0xA_BCDE_1234_5678  ; deposit 0xA_BCDE    into R1 
\end{lstlisting}

loads the constant \texttt{0xA\_BCDE\_1234\_5678} into R1.

\end{iPageDescEntry}