%--------------------------------------------------------------------------------
%
%
%--------------------------------------------------------------------------------
\chapter*{Calling Convention}
\addcontentsline{toc}{chapter}{Calling Convention}

The previous chapters introduced the register-usage conventions and the overall stack layout defined by the runtime architecture. Building on that foundation, this chapter describes the mechanisms used to invoke routines and the conventions that govern control transfer, parameter passing, register preservation, and stack-frame management.

The discussion begins with the simplest form of call: the \emph{local call}, which transfers control between routines within the same module. Local calls illustrate the fundamental steps of the calling convention without the additional considerations required for inter-module calls.

\section*{Calling Sequence}
\addcontentsline{toc}{section}{Calling Sequence}

A routine call proceeds through several well-defined stages. First, the caller evaluates all argument expressions and places the resulting values into the parameter area, either in registers or on the stack as required. After preparing the parameters, the caller saves any registers whose values must remain intact across the call. These caller-saved registers must be preserved by the caller before branching to another routine.

The callee then begins execution at its entry point. It accesses the incoming parameters in the established layout of the parameter area, allocates a stack frame, records the return link, and saves any callee-saved registers it intends to overwrite. This setup ensures that caller and callee operate independently while maintaining a consistent runtime state.

When the callee completes its work, it restores all callee-saved registers, retrieves the return link from the frame marker, deallocates its stack frame, and transfers control back to the caller. This division of responsibility allows efficient cooperation between routines and avoids unnecessary state preservation.

As described earlier, the general-purpose register set is partitioned into caller-save and callee-save regions. A compiler selects registers from these regions as needed, balancing the work of saving and restoring state between the caller and the callee. When possible, redundant saves and restores are eliminated. The runtime architecture distinguishes between two basic categories of calls: \emph{local calls}, which remain within a single module, and \emph{external calls}, which cross module boundaries and require additional handling. The following sections describe these types in detail.

\section*{Local Call}
\addcontentsline{toc}{section}{Local Call}

A local call is the most common form of procedure call. The following code snippet illustrates such a call. The caller loads the arguments, saves any registers the callee might modify, and branches to the callee. After the callee returns, the caller restores its saved registers and continues execution.

\begin{figure}[H]
\centering
\lstset{style=codesnippetstyle}
\begin{lstlisting}
; calling procedure

			...					; load arguments
			...					; save caller saves
			BL target, RL 		; branch to "target", save return
			...					; restore caller saves
\end{lstlisting}
\end{figure}


The callee begins by saving the return link, allocating its stack frame, and saving any registers that must remain unchanged across the call. When the routine completes, it restores the return address, restores callee-saved registers, deallocates the stack frame, and branches back to the caller.

\begin{figure}[H]
\centering

\lstset{style=codesnippetstyle}
\begin{lstlisting}
; called procedure
            
target:     ST RL, -16(SP)	    ; save RL
			ST SP, -8 (SP)      ; save PSP
			LDO SP, size(SP) 	; allocate stack frame
			...					; save callee save
			
			...					; do the work
			
			
			...					; restore callee saves
			LDO SP, -size (SP)  ; deallocate stack frame
			LD RL, -16(SP)		; restore RL
			BV (RL) 			; return to caller
\end{lstlisting}
\end{figure}

This sequence shows all steps that may be required for a local call. Depending on the specific call and the nature of the callee, some steps may be omitted. For example, if the caller does not use any caller-saved registers, no saving is needed before the call. Likewise, a routine with no arguments does not need to evaluate arguments. Also, saving the previous stack pointer is an optional operation.

The callee can also omit steps. A \emph{leaf procedure}, i.e. one that calls no other routines, does not need to save the return address. If it does not use any callee-saved registers, those need not be saved or restored. Finally, if a leaf procedure uses no caller-saved registers, has no local variables, and does not need to preserve temporary registers, it may require no stack frame at all.

The instructions \texttt{B}, \texttt{BR}, and \texttt{BV} are used for local branches. The address computation, whether IA-relative or region-relative, does not modify the region portion of the instruction address. It is always the region of the currently executing code.

\section*{Parameter passing}
\addcontentsline{toc}{section}{Parameter passing}

Procedures may pass arguments to a callee and receive return values. The parameter area is located in the caller's stack frame, below the frame marker.

\begin{figure}[H]
\begin{center}
\begin{tikzpicture}[x=1cm, y=1cm, scale=0.7, transform shape]
        
  	% \draw[help lines, gray!70, dashed] (0,0) grid(16,8);
  	
  	% Rectangle dimensions
  	\def\size{ 10 } 
	\def\width{6}
	\def\leftofs{6}
	\def\rightofs{\leftofs + \width}
	\def\middleofs{\leftofs + \width / 2 }
  	  	
	% Box
	\draw[thick] (\leftofs, 0) -- (\leftofs, \size);
	\draw[thick] (\rightofs, 0) -- (\rightofs, \size);
	\draw[thick] (\leftofs, 0) -- (\rightofs, 0);
	
	\draw[thick, dashed] (\leftofs, \size) -- (\leftofs, \size + 2);
	\draw[thick, dashed] (\rightofs, \size) -- (\rightofs, \size + 2);
	
	\draw[thick, dashed] (\leftofs, 9 ) -- (\rightofs, 9  );
	\draw[thick] (\leftofs, 7 ) -- (\rightofs, 7  );
	\draw[thick] (\leftofs, 6 ) -- (\rightofs, 6  );
	\draw[thick] (\leftofs, 5 ) -- (\rightofs, 5  );
	\draw[thick] (\leftofs, 4 ) -- (\rightofs, 4  );
	\draw[thick] (\leftofs, 3 ) -- (\rightofs, 3  );
	\draw[thick] (\leftofs, 2 ) -- (\rightofs, 2  );
	\draw[thick] (\leftofs, 1 ) -- (\rightofs, 1  );
				  	
  	% Fields
  	\node at ( \middleofs, 6.5 ) { Arg4 };
  	\node at ( \middleofs, 5.5 ) { Arg3 / Ret3};
  	\node at ( \middleofs, 4.5 ) { Arg2 / Ret2 };
  	\node at ( \middleofs, 3.5 ) { Arg1 / Ret1 };
  	\node at ( \middleofs, 2.5 ) { Arg0 / Ret0 };
  	\node at ( \middleofs, 1.5 ) { Return Address };
  	\node at ( \middleofs, 0.5 ) { Previous Frame Pointer (PSP) };
  	
  	%Pointers
  	\node[anchor=east] at (\leftofs - 1, 9) { \texttt{SP - nn} };
  	\node[anchor=east] at (\leftofs - 1, 7) { \texttt{SP - 56} };
  	\node[anchor=east] at (\leftofs - 1, 6) { \texttt{SP - 48} };
  	\node[anchor=east] at (\leftofs - 1, 5) { \texttt{SP - 40} };
  	\node[anchor=east] at (\leftofs - 1, 4) { \texttt{SP - 32} };
  	\node[anchor=east] at (\leftofs - 1, 3) { \texttt{SP - 24} };
  	\node[anchor=east] at (\leftofs - 1, 2) { \texttt{SP - 16} };
  	\node[anchor=east] at (\leftofs - 1, 1) { \texttt{SP - 8} };
  	
  	\draw[->, thick] (\leftofs - 0.5, 0) -- (\leftofs, 0);
  	\node[anchor=east] at (\leftofs - 1, 0) { Stack Pointer (SP) };
  	
  	% Brackets
  	\draw[thick]  [decorate, decoration={brace, amplitude=0.25cm, mirror}] 
  	    (\rightofs + 0.5,6) -- (\rightofs + 0.5,9)
	    node[midway, right=1cm, anchor=west]{\Large Arguments on Stack};
	
  	\draw[thick]  [decorate, decoration={brace, amplitude=0.25cm, mirror}] 
  	    (\rightofs + 0.5,2) -- (\rightofs + 0.5,6)
	    node[midway, right=1cm, anchor=west]{\Large Register Arguments};
	
	\draw[thick]  [decorate, decoration={brace, amplitude=0.25cm, mirror}] 
	    (\rightofs + 0.5,0) -- (\rightofs + 0.5,2)
	    node[midway, right=1cm, anchor=west]{\Large Frame Header};
  	  
\end{tikzpicture}
\end{center}
\caption{Argument Passing}
\label{fig:argument_passing}
\end{figure}

The first four arguments are passed in registers \texttt{ARG0} through \texttt{ARG3}. Nevertheless, the caller's stack frame reserves space for them in the parameter area. Any additional parameters are stored on the stack. Upon return, the callee may supply up to two return values in registers \texttt{RET0} and \texttt{RET1}.

Arguments and return values are addressed relative to the caller's frame pointer. Because the callee may allocate its own stack frame, the size of the callee's frame must be added to the fixed offsets of the arguments in the caller's frame.

\section*{External Calls}
\addcontentsline{toc}{section}{External Calls}

An external call is one that may cross a region boundary. The \texttt{BE} instruction computes the effective address from a general register and an immediate offset encoded in the instruction. Like local branches, this computation produces a 32-bit unsigned address offset. However, unlike local branches, the base register contains the full virtual address, region and offset, not just an offset. The \texttt{BE} instruction is used both to call a target in another region and to return from it. The following code snippets shows a basic external call.

\begin{figure}[H]
\centering
\lstset{style=codesnippetstyle}
\begin{lstlisting}

; calling code

			...					; load arguments
			...					; save caller saves
			LD R1, <target>)    ; obtain the target address    
			BE R1, RL    		; branch to "target", save return
			...					; restore caller saves
\end{lstlisting}
\end{figure}


This snippet illustrates only the external branch. The full runtime convention is more elaborate. When calling an external function, the target address and its associated DP value must be obtained. This information is stored in the \textbf{linkage table}, which is populated with actual values at load time. This indirection allows compiled code to remain unchanged regardless of where the external function is ultimately located.

Because there is no practical way for the callee to determine its correct DP value in a shared-code model, the caller is responsible for setting DP before branching outside its module. The necessary steps of loading the target address, loading the target DP, and performing the external branch are placed into a small piece of code called a \emph{stub}. A stub is generated at compile time and inserted into the module's code. The branch instruction is patched during the linking process.

A generic external call sequence is shown below:

\begin{figure}[H]
\centering
\lstset{style=codesnippetstyle}
\begin{lstlisting}

;calling code

			...
            ...						; load arguments
			...						; save caller saves
			...						; save DP
			B stubF1				; branch to the stub for "F1"
			...						; restore DP
			...						; restore caller saves
			...
			
;stubF1

stubF1:		LD R1, -ofs(DP)		    ; load target address of F1
			LD DP, -ofs + 8(DP)     ; load target DP value
			BE (R1)					; branch to target

\end{lstlisting}
\end{figure}


Before the stub loads the new DP value, the caller must save the current DP. This step is always required, even if the called function does not use DP. The Twin-64 runtime model performs this work in the caller's code before branching to the stub. This division has several benefits. First, the caller can optimize DP saving. For example, by dedicating a register to hold it so that only restoring DP is necessary after external calls. Second, the stub's sequence of loading the target address and DP value is shared for all calls to the same external function. Finally, the target function is unaware of whether it was reached by a local or external call:

\begin{figure}[H]
\centering
\lstset{style=codesnippetstyle}
\begin{lstlisting}

; called code

target:     ST RL, -16(SP)		; save RL
            ST SP,  -8(SP)      ; save PSP
			LDO SP, size(SP)    ; allocate stack frame of <size>
			...					; save callee save regs
			
			...					; do the function work
			
			...					; restore callee saves
			LDO SP, -size(SP)   ; deallocate stack frame
			LD RL, -16(SP)      ; restore RL
			BE (RL) 		    ; branch to return location	
\end{lstlisting}
\end{figure}

Local calls may use \texttt{BV} to return, since the region portion of the address does not change. Exported functions, however, may return either locally or externally, and therefore must use \texttt{BE}. It is recommended to simply always use the \texttt{BE} instruction.

\section*{Linkage Table}
\addcontentsline{toc}{section}{Linkage Table}

The assembler or compiler generates the skeleton of the linkage table. The linkage table resides below the global data area. Entries are addressed using negative DP offsets.

\begin{figure}[H] 
\begin{center} 
\begin{tikzpicture}[x=1cm, y=1cm, scale=0.7, transform shape] 

	\draw[help lines, gray!70, dashed] (0,0) grid(16,16);
	
	\def\dataLeft {4}; 
	\def\codeLeft {9}; 
	\def\width {6}; 
	
	%box
	\node[ tsRectangle, minimum width=6cm, minimum height=14cm, fill=white!50, anchor=west] at ( \dataLeft, 9 ) { };
	
	\draw[thick] (\dataLeft, 12) -- (\dataLeft + \width, 12);
	\draw[thick] (\dataLeft, 8) -- (\dataLeft + \width, 8);
	\draw[thick] (\dataLeft, 2) -- (\dataLeft + \width, 2);
	
	\draw[thick, dashed] (\dataLeft, 0) -- (\dataLeft, 2);
	\draw[thick, dashed] (\dataLeft + \width, 0) -- (\dataLeft + \width, 2);
	
	%left side text
	\node[anchor=east] at (\dataLeft - 1, 16 ) { \Large Data Region }; 
	\draw[<-, thick] (\dataLeft, 16) -- (\dataLeft - 1, 16); 
	
	\draw[<-, thick] (\dataLeft, 12) -- (\dataLeft - 1, 12); 
	\node[anchor=east] at (\dataLeft - 1, 12 ) { \Large DP }; 
	
	\draw[<-, thick] (\dataLeft, 4) -- (\dataLeft - 1, 4); 
	\node[anchor=east] at (\dataLeft - 1, 4 ) { \Large SP }; 

	% global data text    
	\draw[thick]  [decorate, decoration={brace, amplitude=0.25cm, mirror}] 
	    (\dataLeft + 6.5, 8 ) -- (\dataLeft + 6.5, 12 )
	    node[midway, right=1cm, anchor=west]{\Large Global Data};
	    
	% linkage table text    
	\draw[thick]  [decorate, decoration={brace, amplitude=0.25cm, mirror}] 
	    (\dataLeft + 6.5, 12 ) -- (\dataLeft + 6.5, 16 )
	    node[midway, right=1cm, anchor=west]{\Large Linkage Table};
	    
	% stack area text    
	\draw[thick]  [decorate, decoration={brace, amplitude=0.25cm, mirror}] 
	 	(\dataLeft + 6.5, 2 ) -- (\dataLeft + 6.5, 8 )
	    node[midway, right=1cm, anchor=west]{ \Large Stack Area};
	    
	% module A externals        
	\node[ tsRectangle, minimum width=6cm, minimum height=1cm, fill=gray!40, anchor=west] at ( \dataLeft, 14 ) { }; 
	\node[anchor=west] at (\dataLeft + 0.25, 14) { \small Module A Externals};
	% module B externals
	\node[ tsRectangle, minimum width=6cm, minimum height=1cm, fill=gray!40, anchor=west] at ( \dataLeft, 13 ) { }; 
	\node[anchor=west] at (\dataLeft + 0.25, 13) { \small Module B Externals};
	
	% module A globals
	\node[ tsRectangle, minimum width=6cm, minimum height=1cm, fill=gray!40, anchor=west] at ( \dataLeft, 11 ) { }; 
	\node[anchor=west] at (\dataLeft + 0.25, 11) { \small Module A Global Data};
	
	% module B globals
	\node[ tsRectangle, minimum width=6cm, minimum height=1cm, fill=gray!40, anchor=west] at ( \dataLeft, 10 ) { }; 
	\node[anchor=west] at (\dataLeft + 0.25, 10) { \small Module B Global Data};
	
	% stack frame
	\node[ tsRectangle, minimum width=6cm, minimum height=1cm, fill=gray!40, anchor=west] at ( \dataLeft, 4.5 ) { }; 
	\node[anchor=west] at (\dataLeft + 0.25, 4.5) { \small Stack Frame};
	
	% rest
	\filldraw ( \dataLeft + 2.5, 1) circle (0.5mm);
  	\filldraw ( \dataLeft + 3, 1) circle (0.5mm);
  	\filldraw ( \dataLeft + 3.5, 1) circle (0.5mm);
	
\end{tikzpicture} 
\end{center} 
\caption{Linkage Table} 
\label{fig: Linkage Table} 
\end{figure}

The figure illustrates two modules within a task's global data region. All module global data is addressed as DP plus an offset, while linkage-table entries are addressed as DP minus an offset. The compiler constructs the linkage table during compilation. The linkage table contains per external library several entries. 

The DP value entry starts the linkage subtable for a particular library. 

\begin{center}
    \drawInstrFormatSixtyFour
       {0,3,16}
        {0, DP value of external Function Library}
\end{center}

The function target entry contains the target address of the function.

\begin{center}
    \drawInstrFormatSixtyFour
       {0,3,16}
        {0, Virtual Address of external Function}
\end{center}


\section*{Function Labels}
\addcontentsline{toc}{section}{Function Labels}

A function label is a reference to a function which is resolved at calling time. Therefore each function that is accessed via a function label has an entry in the linkage table. The function label itself is then the index into the linkage table. 

??? the entry is two pieces: DP index and function index. To be defined...
\begin{center}
    \drawInstrFormatSixtyFour
       {0,8,12,16}
        {0, DP index, Function index}
\end{center}


\begin{figure}[H]
\centering
\lstset{style=codesnippetstyle}
\begin{lstlisting}

; example
			...					; 
			...					; 
		
\end{lstlisting}
\end{figure}



\section*{Privilege Level Changes}
\addcontentsline{toc}{section}{Privilege Level Changes}


\begin{figure}[H]
\centering
\lstset{style=codesnippetstyle}
\begin{lstlisting}

; example
			...					; 
			...					; 
		
\end{lstlisting}
\end{figure}


\section*{Design Rationale}
\addcontentsline{toc}{section}{Design Rationale}

There are many possible ways to organize modules and their external linkages. Classic CISC architectures provided simple call instructions that implicitly performed all the required steps such as address computation, access-rights checking, and memory references needed to resolve external linkages.

RISC designs, in contrast, expose only simple branch and jump instructions and rely on explicit caller–-callee code sequences to implement calling conventions. Because compilers require a unified and predictable calling sequence, the additional work needed for external calls is typically done via import and export stubs.

Twin-64 follows a similar model. External calls use an import stub, which is responsible for locating the target function and obtaining the external library’s DP value before issuing the external branch instruction. The caller saves its own DP value and restores it upon return.

This approach adds two instructions at each external call site, but provides a key benefit: it avoids two additional branches, one from the import stub to the callee and one back from the import stub to the caller. It also centralizes DP handling. Because the stub does not manipulate the caller’s DP value, the caller has full flexibility. It may save DP to memory, move it to another register, or even save it once and reload it after each external call, depending on optimization goals.

Thus an external call requires only one additional branch in the calling sequence, and the code expansion for saving and restoring the caller’s DP is modest. The callee is completely unaware of the caller’s region, so no export stub is required.

An additional advantage is that the compiler does not need to generate any special code for functions that may be called externally. This reduces complexity and places less burden on the compiler designer.
